\subsection{January}

\subsubsection{January 1st}
Today I learned that the Ackermann function can be expressed using our generalized primitive recursion over $\NN.$ Typically the Ackermann function isn't primitive recursive because we only allow recursion over $\NN,$ and the Ackermann function is more or less designed to eventually overcome however much nesting we do. However, in type theory we may recurse over any type, not just $\NN,$ which is apparently significantly more power. The idea now is that we can do recursion over recursive functions (not just over $\NN$), which should give us enough power.

Anyways, we are interested in defining $\op{ack}:\NN\to\NN\to\NN$ satisfying
\begin{align*}
    \op{ack}(0,n) &\equiv \op{succ}(n), \\
    \op{ack}(\op{succ}(m),0) &\equiv \op{ack}(m,1), \\
    \op{ack}(\op{succ}(m),\op{succ}(n)) &\equiv \op{ack}(m,\op{ack}(\op{succ}(m),n)).
\end{align*}
The last two equations can be read as
\begin{align*}
    \op{ack}(\op{succ}(m))(0) &\equiv\op{ack}(m)(1), \\
    \op{ack}(\op{succ}(m))(\op{succ}(n)) &\equiv\op{ack}(m)\big(\op{ack}(\op{succ}(m))(n)\big).
\end{align*}
However, we can read this as just recursively defining $\op{ack}(\op{succ}(m))$ given $\op{ack}(m).$ Namely, we see we can write
\[\op{ack}(\op{succ}(m))\equiv\op{rec}_\NN(\NN,\op{ack}(m)(1),\lambda x.\lambda n.\op{ack}(m)(n))\]
from the definition of the recursor.

Going further, we note that we can read the above as a step function for functions $\NN\to\NN$, for which
\[\op{ack}(0)\equiv\op{succ}\]
is the base case. We want our step function $\NN\to(\NN\to\NN)\to(\NN\to\NN)$ to take $m$ and the previous $\op{ack}(m)$ and output $\op{ack}(\op{succ}(m)),$ so we note
\[\lambda y.\lambda f.\op{rec}_\NN(\NN,f(1),\lambda x.\lambda n.f(n))\]
will work. We see this because applying $(m,\op{ack}(m))$ to the above does indeed give $\op{ack}(\op{succ}(m))$ from our work. Anywyas, this means we can recurse to write
\[\boxed{\op{ack}\equiv\op{rec}_\NN\big(\NN\times\NN,\op{succ},\lambda y.\lambda f.\op{rec}_\NN(\NN,f(1),\lambda x.\lambda n.f(n))\big)}.\]
This is what we wanted.

This proof kind of weirds me out. I don't think that allowing generalized recursion over types makes us Turing-complete, for all the functions we make must be defined on all inputs, but we certainly have more power than I'm used to having. We have exhibited more power than $\texttt{LOOP},$ and I would be interested in seeing a computable function we can't make with fancy recursion. Of course, it's not even clear to me how we could formally show a function can't be defined with our recursion.

\subsubsection{January 2nd}
Today I learned some properties of equality. The one I found most remarkable is how type theory deals with the ``substitution'' property of equality. Roughly speaking, we want to show that for $x,y:A,$ given an expression $\Phi,$ we know $x=y$ implies $\Phi_x=\Phi_y,$ where $\Phi_\bullet$ replaces needed instances with $\bullet.$

The way we can talk about expressions like $\Phi$ is with $\lambda$ calculus. Namely, our formalization will show that
\[(x=_Ay)\to(f(x)=_Bf(y))\]
for any function $f:A\to B.$ This means that for $f\equiv\lambda z.\Phi,$ judgemental equality lets us write
\[(x=_Ay)\to(\Phi_x=_B\Phi_y).\]
There is some technicality in that we might not want to assume that $\Phi_x$ and $\Phi_y$ are the same type, but this is good enough for me. Anyways, this reduction is basically saying that the substitution property of equality is (roughly) the same idea as being able to ``do the same thing to both sides of the equal sign.'' I find this somewhat cute.

It remains to inhabit
\[\prod_{x,y:A}\prod_{p:(x=_Ay)}f(x)=_Bf(y).\]
For this we use path induction. Let our family be $C:\prod_{x,y:A}(x=y)\to\mathcal U$ by $C(x,y,p):\equiv f(x)=_Bf(t)$ so that we want to exhibit $\prod_{x,y:A}\prod_{p:(x=y)}C(x,y,p).$

It remains to show the conditions of path induction on our family $C.$ The actual content of the prove lives in the case where $x\equiv y,$ but here $f(x)\equiv f(y),$ meaning $\op{refl}_{f(x)}:f(x)=f(y)$ can witness. Thus, we may set
\[c(x):\equiv\op{refl}_{f(x)}:\prod_{x:A}C(x,x,\op{refl}_x).\]
This finishes the proof by path induction.

Path induction in general makes quick work of equality statements. In the above, what feels so weird is that it's not at all obvious how we would go about constructing a witness of $f(x)=_Bf(y)$ from $x=_Ay,$ but nonetheless we know that the witness exists. Namely, we've only shown the witness only for the most trivial case, in which they are judgementally equal.

This nonconstructivity feels a bit off in type theory, but it is certainly very powerful. The proof may be obnoxious to write out, but we could just have well said the following: ``By path induction, it suffices to show that $(x=_Ay)\to f(x)=_Bf(y)$ in the case $x\equiv y$ and $\op{refl}_x$ is our input. But then $f(x)\equiv f(y),$ so we may take $\op{refl}_x\to\op{refl}_{f(x)},$ which finishes.''

\subsubsection{January 3rd}
Today I learned the details of the Eckmann-Hamilton theorem in type theory. The main ingredient is horizontal composition, defined with whiskering. Horizontal composition is defined over homotopies of homotopies, in the following setup.
\begin{center}
    \begin{tikzcd}
        a \arrow[rr, "p", bend left, shift left] \arrow[rr, "q"', bend right, shift right] & \Downarrow\alpha & b \arrow[rr, "r", bend left, shift left] \arrow[rr, "s"', bend right, shift right] & \Downarrow\beta & c
    \end{tikzcd}
\end{center}
Here, $a,b,c:A$ where $p,q:a=_Ab$ and $r,s:b=_Ac,$ with even $\alpha:p=_{a=b}q$ and $\beta:r=_{b=c}s.$ The idea is that we should be able to exhibit something like \[\alpha\star\beta:p\cdot r=_{a=c}q\cdot s\]
by imagining moving $p\to q$ by $\alpha$ and $r\to s$ by $\beta.$ This is horizontal composition.

To do this, we move the homotopies one at a time, which is called whiskering. Namely, we exhibit
\[\alpha\cdot_{r}r:p\cdot r=q\cdot r,\qquad q\cdot_{\ell}\beta:q\cdot r=q\cdot s.\]
Imagine $\alpha\cdot_rr$ as fixing the $r$ight whisker $r$ and dragging $p\to q$ by $\alpha$ and then $q\cdot_\ell\beta$ fixes the $\ell$eft whisker and then drags $r\to s$ by $\beta.$ This is doing what we wanted horizontal composition to do. Observe that composing will witness
\[p\cdot r\stackrel{\alpha\cdot_rr}=q\cdot r\stackrel{q\cdot_\ell\beta}=q\cdot s,\]
which is exactly what we wanted. So it remains to exhibit $\cdot_r$ and $\cdot_\ell.$ I'll do one of these formally; the other is similar.

We define $\alpha\cdot_rr$ by induction on $r$: more precisely, with $b$ fixed (it's involved in $\alpha$'s information), we're going to let $r$ vary with $c,$ creating a path induction based at $b.$ That is, we think about constructing like
\[(\alpha\cdot_r-):\prod_{c:A}\prod_{r:(b=c)}p\cdot r=q\cdot r.\]
Our family for the based path induction will be
\[C:\equiv\lambda(c:A).\lambda(r:b=c).p\cdot r=q\cdot r,\]
so it remains to exhibit $c:C(b,\refl_b)\equiv p\cdot\refl_b=q\cdot\refl_b$ to construct our function. For this, we note that $p=p\cdot\refl_b$ is inhabited by some $\op{ru}_p,$ and similar for $q=q\cdot\refl_q$ inhabited by $\op{lu}_p.$ This implies we can write
\[p\cdot\refl_b\stackrel{\op{ru}_p^{-1}}=p\stackrel\alpha=q\stackrel{\op{ru}_q}=q\cdot\refl_b,\]
so $c:\equiv\op{ru}_p^{-1}\cdot\alpha\cdot\op{ru}_q:p\cdot\refl_b=q\cdot\refl_b.$ So path induction will define $\alpha\cdot_rr$ with base case $\alpha\cdot_r\refl_b\equiv\op{ru}_p^{-1}\cdot\alpha\cdot\op{ru}_q.$

The case with $q\cdot_\ell\beta$ is pretty much the same, which I'll outline. Again, we want to let $q$ vary in our induction, but we can't move $b,$ so we do a based induction on $b$ where $a$ varies. Then the base case is where $a\equiv b$ and $q\equiv\refl_b,$ in which case we need to witness $\refl_b\cdot r=\refl_b\cdot s.$ But as before, we write
\[\refl_b\cdot r\stackrel{\op{lu}_r^{-1}}=r\stackrel\beta=s\stackrel{\op{lu}_s}=s\cdot\refl_b,\]
so $\op{lu}_r^{-1}\cdot\beta\cdot\op{lu}_s$ will witness.

Note that $s$ does not matter in our construction of $\alpha\cdot_rr,$ so we just have well have $\alpha\cdot_rs:p\cdot s=q\cdot s$ by just plugging in $s$ to the same function $(\alpha\cdot_r-)$ we constructed. Similarly, we also have a $p\cdot_\ell\beta:p\cdot r=p\cdot s$ by plugging in $p$ to our $(-\cdot_\ell\beta).$ Composing again, we see
\[p\cdot r\stackrel{p\cdot_\ell\beta}=p\cdot s\stackrel{\alpha\cdot_rs}=q\cdot s,\]
so we could have defined $\alpha\star'\beta:p\cdot r=q\cdot s$ by $(p\cdot_\ell\beta)\cdot(\alpha\cdot_rs).$ In other words, the order of which whisker we drag down shouldn't matter.

In order to sleep better at night, we had better have $\alpha\star\beta=\alpha\star'\beta.$ Showing this requires many layers of induction; I do it formally for practice. Expanding a bit, we want to witness
\[f_1:\prod_{(a,b,c:A)}\prod_{(p,q:a=b)}\prod_{(\alpha:p=q)}\prod_{(r,s:b=c)}\prod_{(\beta:r=s)}(\alpha\cdot_rr)\cdot(q\cdot_\ell\beta)=(p\cdot_\ell\beta)\cdot(\alpha\cdot_rs).\]
Obviously, we use path induction. With nowhere else to start, we start from the inside. That is, we read this as wanting to exhibit
\[f_1(a,b,c,p,q,\alpha):\prod_{(r,s:b=c)}\prod_{(\beta:r=s)}(\alpha\cdot_rr)\cdot(q\cdot_\ell\beta)=(p\cdot_\ell\beta)\cdot(\alpha\cdot_rs).\]
We work with the family
\[C_1(a,b,c,p,q,\alpha)(r,s,\beta)\equiv(\alpha\cdot_rr)\cdot(q\cdot_\ell\beta)=(p\cdot_\ell\beta)\cdot(\alpha\cdot_rs)\]
so that it suffices to provide $f_1$ only in the case $C_1(r,r,\refl_r).$ That is, it suffices to exhibit
\[f_2:\prod_{(a,b,c:A)}\prod_{(p,q:a=b)}\prod_{(\alpha:p=q)}\prod_{(r:b=c)}(\alpha\cdot_rr)\cdot(q\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\alpha\cdot_rr).\]
Continuing with the path induction, we read this as wanting to exhibit
\[f_2(a,b,c):\prod_{(p,q:a=b)}\prod_{(\alpha:p=q)}\prod_{(r:b=c)}(\alpha\cdot_rr)\cdot(q\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\alpha\cdot_rr)\]
to work with the family
\[C_2(a,b,c)\equiv\prod_{(p,q:a=b)}\prod_{(\alpha:p=q)}\prod_{(r:b=c)}(\alpha\cdot_rr)\cdot(q\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\alpha\cdot_rr).\]
By path induction, it now only suffices to provide $f_2$ in the case $C_2(p,p,\refl_p).$p That is, it suffices to exhibit
\[f_3:\prod_{(a,b,c:A)}\prod_{(p:a=b)}\prod_{(r:b=c)}(\refl_p\cdot_rr)\cdot(p\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\refl_p\cdot_rr).\]
Had we defined whiskering by induction on $\alpha$ and $\beta,$ we could finish here with some judgemental equalities. Alas, we have more induction to do. Swapping, we see it's sufficient to exhibit
\[f_3':\prod_{(a,b:A)}\prod_{(p:a=b)}\prod_{(c:A)}\prod_{(r:b=c)}(\refl_p\cdot_rr)\cdot(p\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\refl_p\cdot_rr).\]
For this, we use induction, our family being
\[C_3(a,b,p)\equiv\prod_{(c:A)}\prod_{(r:b=c)}(\refl_p\cdot_rr)\cdot(p\cdot_\ell\refl_r)=(p\cdot_\ell\refl_r)\cdot(\refl_p\cdot_rr)\]
so that path induction says it's enough to provide $f_3'$ in the case $C_3(a,a,\refl_a).$ That is, it suffices to exhibit
\[f_4:\prod_{(a,c:A)}\prod_{(r:b=c)}(\refl_{\refl_a}\cdot_rr)\cdot(\refl_a\cdot_\ell\refl_r)=(\refl_a\cdot_\ell\refl_r)\cdot(\refl_{\refl_a}\cdot_rr)\]
For this, we use induction, our family being
\[C_4(a,c,r)\equiv(\refl_{\refl_a}\cdot_rr)\cdot(\refl_a\cdot_\ell\refl_r)=(\refl_a\cdot_\ell\refl_r)\cdot(\refl_{\refl_a}\cdot_rr)\]
so that path induction says it's enough provide $f_r$ in the case $C_4(a,a,\refl_a).$ That is, ut suffices to exhibit
\[f_5:\prod_{(a:A)}(\refl_{\refl_a}\cdot_r\refl_a)\cdot(\refl_a\cdot_\ell\refl_{\refl_a})=(\refl_a\cdot_\ell\refl_{\refl_a})\cdot(\refl_{\refl_a}\cdot_r\refl_a).\]

Now for the content of the proof; we are almost done. We take $a\equiv b\equiv c$ and $p\equiv q\equiv r\equiv s\equiv\refl_a$ so that $\refl_a\cdot\refl_a\equiv\refl_a,$ implying
\[\op{ru}^{-1}_{\refl_a}\equiv\op{ru}_{\refl_a}\equiv\op{lu}_{\refl_a}\equiv\op{lu}^{-1}_{\refl_a}\equiv\refl_{\refl_a}.\]
In particular,
\[\alpha\cdot_r\refl_a\equiv\op{ru}_{\refl_a}^{-1}\cdot\alpha\cdot\op{ru}_{\refl_a}\equiv\refl_{\refl_a}\cdot\alpha\cdot\refl_{\refl_a}\equiv\alpha,\]
and
\[\refl_a\cdot_\ell\beta\equiv\op{lu}_{\refl_a}^{-1}\cdot\beta\cdot\op{lu}_{\refl_a}\equiv\refl_{\refl_a}\cdot\beta\cdot\refl_{\refl_a}\equiv\beta.\]
This means that $(\refl_{\refl_a}\cdot_r\refl_a)\cdot(\refl_a\cdot_\ell\refl_{\refl_a})\equiv(\refl_a\cdot_\ell\refl_{\refl_a})\cdot(\refl_{\refl_a}\cdot_r\refl_a)\equiv\refl_{\refl_a},$ so we may just do
\[f_5\equiv\lambda a.\refl_{\refl_{\refl_a}}.\]
And with that, our induction is complete.

Anyways, the point of this is to say that in the event $a\equiv b\equiv c$ and $p\equiv q\equiv r\equiv s$ so that $\alpha,\beta\in\Omega^2(A,a),$ we still have
\[\alpha\star\beta=\alpha\star'\beta.\]
However, expanding these both out tells us
\[\alpha\cdot\beta\equiv(\alpha\cdot_r\refl_a)\cdot(\refl_a\cdot_\ell\beta)\equiv\alpha\star\beta\]
while
\[\alpha\star'\beta\equiv(\refl_a\cdot_\ell\beta)\cdot(\alpha\cdot_r\refl_a)\equiv\beta\cdot\alpha.\]
In particular, $\Omega^2(A,a)$ is abelian. This is nice, so we call it here.

\subsubsection{January 4th}
Today I learned about type-theoretic homotopies between (dependent) functions. Fix a type $A$ with family $P:A\to\UU.$ Then for dependent functions $f,g:\prod_{(x:A)}P(x),$ we define the type of homotopies as
\[f\sim g\equiv\prod_{x:A}(f(x)=g(x)).\]
Essentially, we're saying that the two functions are ``the same'' (homotopically equivalent) if they are pairwise ``identical.'' Topologically, we can view this as saying that we have a path from each $f(x)$ to $g(x),$ so we can visualize the corresponding deformation/homotopy from $g$ to $g$ along each input.
\begin{center}
    \begin{asy}
        import graph;
        unitsize(1cm);
        real x(real t)
        {
            return (sin(4*pi*t)-1)/13+(cos(7*t)-1)/11-1/3-1/10;
        }
        real x1(real t)
        {
            return x(t);
        }
        real x2(real t)
        {
            return 2+x(t+10.987);
        }
        real y(real t)
        {
            return 4*t;
        }
        draw(graph(x1,y,0,1));
        draw(graph(x2,y,0,1));
        real t1(real t)
        {
            return t;
        }
        real t2(real t)
        {
            return t*(1-sin(5)/5) + sin(5*t)/5;
        }
        int N=8;
        pair p1, p2;
        for(int i = 0; i <= N; ++i)
        {
            p1 = ( x1(t1(i/N)) , y(t1(i/N)) );
            p2 = ( x2(t2(i/N)) , y(t2(i/N)) );
            draw(p1--p2, arrow=EndArrow, dashed);
            if(i==2)
            {
                dot("$f(x)$",p1,W);
                dot("$g(x)$",p2,E);
            }
        }
    \end{asy}
\end{center}
In this way, we can imagine deforming (the graph of) each $f(x)$ and $g(x)$ along the given paths, which is our homotopy. In practice, the fact that this is a dependent function makes this visual not make as much sense---the $f(x)$ and $g(x)$ can live in vastly different universes.

Anyways, homotopy behaves the way we want it to. For example, it forms an equivalence class of functions. Reflexive means that we can witness
\[\prod_{f:\prod_{(x:A)}P(x)}(f\sim f)\equiv\prod_{\left(f:\prod_{(x:A)}P(x)\right)}\prod_{(x:A)}(f(x)=f(x)),\]
for which $\lambda f.\lambda x.\refl_{f(x)}$ will work because $\refl_{f(x)}:f(x)=f(x).$ Symmetric means that we can witness
\[\prod_{f,g:\prod_{(x:A)}P(x)}(f\sim g)\to(g\sim f)\equiv\prod_{\left(f,g:\prod_{(x:A)}P(x)\right)}\left(\prod_{x:A}f(x)=g(x)\right)\to\left(\prod_{x:A}g(x)=f(x)\right),\]
for which $\lambda f.\lambda g.\lambda(H:f\sim g).\lambda(x:A).H(x)^{-1}$ will work because $H(x)^{-1}:g(x)=f(x).$ Transitivity means that we can witness
\[\prod_{f,g,h:\prod_{(x:A)}P(x)}(f\sim g)\to(g\sim h)\to(f\sim h),\]
which is
\[\prod_{\left(f,g,h:\prod_{(x:A)}P(x)\right)}\left(\prod_{x:A}f(x)=g(x)\right)\to\left(\prod_{x:A}g(x)=h(x)\right)\to\left(\prod_{x:A}f(x)=h(x)\right),\]
for which $\lambda f.\lambda g.\lambda h.\lambda(H_1:f\sim g).\lambda(H_2:g\sim h).\lambda(x:A).H_1(x)\cdot H_2(x)$ will work because $H_1(x)\cdot H_2(x):f(x)=h(x).$

\subsubsection{January 5th}
Today I learned about equivalence of types. Fix $A$ and $B$ types. For $f:A\to B,$ we say that $f$ has a ``quasi-inverse'' $g:B\to A$ if
\[(f\circ g\sim\op{id}_A)\times(g\circ f\sim\op{id}_B).\]
That is,
\[\op{qinv}(f):\equiv\sum_{(g:B\to A)}(f\circ g\sim\op{id}_B)\times(g\circ f\sim\op{id}_A).\]
In words, $f\circ g$ and $g\circ f$ are homotopically the identity, suggesting we have a homeomorphism from $A$ to $B,$ giving us an ``equivalence.''

In reality, $\op{qinv}$ isn't well-behaved, so there's a different type $\op{isequiv}(f)$ which just witnesses $f$ being an equivalence of $A$ and $B.$ What's important is that $\op{isequiv}(f)$ is a bit better-behaved (all its elements are propositionally equal), but $\op{isequiv}(f)$ is inhabited if and only if $\op{qinv}(f)$ is inhabited. Then we define
\[A\simeq B:\equiv\sum_{f:A\to B}\op{isequiv}(f).\]
In accordance to constructive mathematics, we have $A$ is equivalent to $B$ if and only if we can construct the equivalence $f$ and the witness of its equivalence $\op{isequiv}(f).$

Equivalence behaves how we'd like; e.g., it's an equivalence relation over types. Reflexivity $A\simeq A$ is witnessed by $\op{id}_A,$ which has quasi-inverse $\op{id}_A.$ Indeed,
\[\op{id}_A\circ\op{id}_A\sim\op{id}_A\equiv\op{id}_A\sim\op{id}_A\equiv\prod_{x:A}(\op{id}_A(x)=\op{id}_A(x))\]
is witnessed by $\lambda x.\op{refl}_x.$ So $(\op{id}_A,\lambda x.\refl_x,\lambda x.\refl_x):\op{qinv}(\op{id}_A),$ which implies $\op{isequiv}(\op{id}_A)$ is inhabited, so $A\simeq B$ is inhabited.

Symmetry that $(A\simeq B)\to(B\simeq A)$ holds because a function is the quasi-inverse of any of its quasi inverses. Namely, $A\simeq B$ inhabited gives us $f:A\to B$ and $(f^{-1},\alpha,\beta):\op{qinv}(f).$ Well, by construction
\[\begin{cases}
    \beta:f^{-1}\circ f\sim\op{id}_A, \\
    \alpha:f\circ f^{-1}\sim\op{id}_B.
\end{cases}\]
This means $(f,\beta,\alpha):\op{qinv}(f^{-1}).$ Thus, with $f^{-1}:B\to A$ having $\op{qinv}(f^{-1})$ inhabited means $\op{isequiv}(f^{-1})$ is inhabited, so $B\simeq A$ is inhabited, as desired.

Transitivity that $(A\simeq B)\to(B\simeq C)\to(A\simeq C)$ holds because the composition of two functions with a quasi-inverse also has a quasi-inverse. Again, if we are told $A\simeq B$ and $B\simeq C$ are inhabited, then we can build $f:A\to B$ and $g:B\to C$ with $(f^{-1},\alpha_f,\beta_f):\op{qinv}(f)$ and $(g^{-1},\alpha_g,\beta_g):\op{qinv}(g).$ We want to show that $g\circ f:A\to C$ has a quasi-inverse, and we claim that $f^{-1}\circ g^{-1}:C\to A$ is our inverse function. To start, note for $c:C,$
\begin{align*}
    (g\circ f)\circ(f^{-1}\circ g^{-1})(c) &\equiv g\big((f\circ f^{-1})(g^{-1}(c))\big) \\
    &= g\big(\op{id}_B\left(g^{-1}(c)\right)\big) \tag{by $\alpha_f\left(g^{-1}(c)\right)$} \\
    &\equiv g(g^{-1}(c)) \\
    &= \op{id}_C(c). \tag{by $\alpha_g(c)$}
\end{align*}
It follows $\alpha_{gf}\equiv\lambda c.\alpha_f\left(g^{-1}(c)\right)\cdot\alpha_g(c):(g\circ f)\circ(f^{-1}\circ g^{-1})\sim\op{id}_C.$ Similarly, for $a:A,$ we have
\begin{align*}
    (f^{-1}\circ g^{-1})\circ(g\circ f)(a) &\equiv f^{-1}\big((g^{-1}\circ g)(f(a))\big) \\
    &= f^{-1}\big(\op{id}_B(f(a))\big) \tag{by $\beta_g(f(a))$} \\
    &\equiv f^{-1}(f(a)) \\
    &= \op{id}_A(a). \tag{by $\beta_f(a)$}
\end{align*}
It follows $\beta_{gf}\equiv\lambda a.\beta_g(f(a))\cdot\beta_f(a):(f^{-1}\circ g^{-1})\circ(g\circ f)\sim\op{id}_A.$ Bringing our witnesses together, we see
\[\left(f^{-1}\circ g^{-1},\alpha_{fg},\beta_{fg}\right):\op{qinv}(g\circ f).\]
With $\op{qinv}(g\circ f)$ inhabited, we see $\op{isequiv}(g\circ f)$ is inhabited, so $A\simeq C$ is inhabited, as desired.

\subsubsection{January 6th}
Today I learned that equality types of Cartesian products are well-behaved in type theory. Essentially, we have the equivalences
\[\prod_{x,x':A\times B}(x=x')\simeq(\op{pr}_1x=\op{pr}_1x')\times(\op{pr}_2x=\op{pr}_2x').\]
Note I may start repressing the parentheses around single-variable functions like $\op{pr}_1$ when little confusion is possible.

In one direction, we can just run $\op{pr}_1$ and $\op{pr}_2$ to $x=y$ to generate functions
\[\begin{cases}
    \op{ap}_{\op{pr}_1}:(x=x')\to(\op{pr}_1x=\op{pr}_1x'), \\
    \op{ap}_{\op{pr}_2}:(x=x')\to(\op{pr}_2x=\op{pr}_2x').
\end{cases}\]
Then induction for products will give 
\[f(x,x')\equiv\lambda p.\big(\op{ap}_{\op{pr}_1}p,\,\op{ap}_{\op{pr}_2}p\big):(x=x')\to(\op{pr}_1x=\op{pr}_1x')\times(\op{pr}_2x=\op{pr}_2x')\]
because ordered pairs are primitive.

In the other direction, we note that we can use some induction to construct our function. We need to exhibit
\[g:\prod_{x,x':A\times B}(\op{pr}_1x=\op{pr}_1x')\times(\op{pr}_2x=\op{pr}_2x')\to(x=x').\]
By induction on the product, it suffices to exhibit a curried function in
\[\prod_{x,x':A\times B}(\op{pr}_1x=\op{pr}_1x')\to(\op{pr}_2x=\op{pr}_2x')\to(x=x').\]
Now inducting on the product $A\times B,$ we can let $x\equiv(a,b)$ and $x'\equiv(a',b'),$ meaning we have to exhibit
\[\prod_{(a,a':A)}\prod_{(b,b':B)}(a=a')\to(b=b')\to((a,b)=(a',b')).\]
Only at this point we can read this as (nested) path induction, after some swapping. Namely, we exhibit
\[\prod_{(a,a':A)}\prod_{(p:a=a')}\prod_{(b,b':B)}\prod_{(q:b=b')}(a,b)=(a',b').\]
By path induction (family $C(a,a',p)\equiv\prod_{(b,b':B)}\prod_{(q:b=b')}(a,b)=(a',b')$), it suffices to take $a\equiv a'$ and $p\equiv\refl_a.$ So it remains to exhibit
\[\prod_{(a:A)}\prod_{(b,b':B)}\prod_{(q:b=b')}(a,b)=(a,b').\]
By path induction (family $D(b,b',q)\equiv(a,b)=(a,b')$), it suffices to take $b\equiv b'$ and $q\equiv\refl_b.$ But then we see
\[\lambda a.\lambda b.\refl_{(a,b)}:\prod_{(a:A)}\prod_{b:B}(a,b)=(a,b),\]
which is what we wanted.

So we have somewhat natural mappings in both directions. I guess it remains to show that $g(x,x'):\op{qinv}(f(x,x')),$ which will imply $\op{isequiv}(f(x,x'))$ is inhabited, verifying that $f(x,x')$ witnesses the equivalences
\[(x=x')\simeq(\op{pr}_1x=\op{pr}_1x')\times(\op{pr}_2x=\op{pr}_2x').\]
The proof that $g(x,x'):\op{qinv}(f(x,x'))$ is a couple of somewhat involved inductions. I won't do this in formality, for my own sanity.

In one direction, we show $f(x,x')\circ g(x,x')\sim\op{id}_\square,$ which means we have $r:(\op{pr}_1x=\op{pr}_1x')\times(\op{pr}_1x=\op{pr}_1x')$ and need to show
\[(f(x,x')\circ g(x,x'))(r)=r.\]
By induction on $x,x':A\times B,$ we may assert $x\equiv(a,b)$ and $x'\equiv(a',b')$ so that $r:(a=a')\times(b=b').$ It's not clear to me why we can't start with an induction on $r,$ but an induction on $r$ now lets us assert $r\equiv(p,q)$ with $p:a=a'$ and $q:b=b'.$ Now we induct on $p$ and $q$ to let us say $p\equiv\refl_a$ and $q\equiv\refl_b.$ But then
\[(f(x,x)\circ g(x,x))((\refl_a,\refl_b))\equiv f(x,x)(\refl_{(a,b)})\equiv(\refl_a,\refl_b),\]
so reflexivity witnesses here.

In the other direction, we show $g(x,x')\circ f(x,x')\sim\op{id}_\square,$ which means we have $r:x=x'$ and need to show
\[(g(x,x')\circ f(x,x'))(r)=r.\]
By induction on $r,$ we take $x\equiv x'$ and $r\equiv\refl_x.$ Then induction on $x:A\times B$ lets us say $x\equiv(a,b),$ which implies
\[(g(x,x)\circ f(x,x))(\refl_x)\equiv g(x,x)((\refl_a,\refl_b))\equiv\refl_{(a,b)},\]
so again reflexivity witnesses.

\subsubsection{January 7th}
Today I learned the proof of the Newton polygon theorem. I apologize, but I am too tired to make pictures today. Fix $K$ a field equipped with a (nonarchimedean) multiplicative valuation $\nu$; that is, $\nu(x+y)\ge\min\{\nu(x),\nu(y)\}.$ Additionally, we fix a polynomial $f(x)=a_0x^0+\cdots+a_nx^n\in K[x]$ with $a_0,a_n\ne0.$ (We can have $a_0=0,$ but it makes the argument more obnoxious, and $a_0=0$ tells us $f$ has a root at $0$ anyways.) Then we build the Newton polygon from the points
\[(0,\nu(a_0)),\,(1,\nu(a_1)),\,\ldots,\,(n,\nu(a_n)).\]
Namely, we take
\[(s_0,\nu(a_{s_0})),\,(s_1,\nu(a_{s_1})),\,\ldots,\,(s_r,\nu(a_{s_r}))\]
to be the points defining the lower convex hull of our points; note $s_0=0$ and $s_r=n.$ Then the claim is that exactly $s_{k+1}-s_k$ roots of $f$ in $\overline K$ have valuation
\[-\frac{\nu(a_{s_{k+1}})-\nu(a_{s_k})}{a_{s_{k+1}}-a_{s_k}}.\]
That is, the horizontal length of a segment of the Newton polygon gives the number of solutions with valuation equal to the signed slope of that segment. Total horizontal length is $n,$ so this will describe all of our roots, which is quite remarkable.

Let's prove this. It's quite nice. For the moment forget about the Newton polygon; we're going to manually construct it from right to left. Note that we can take $a_n=1$ by dividing $f$ by $a_n.$ This merely shifts down our Newton polygon to hit $(n,0)$ but does not change any of our slopes. (Explicitly, each $\nu(a_\bullet)$ gets $\nu(a_n)$ subtracted, which doesn't affect the difference between the $\nu(a_\bullet)$s.) Now label our roots $\alpha_0,\,\ldots,\,\alpha_{n-1}\in\overline K$ so that
\[\nu(\alpha_{s_k})=\nu(\alpha_{s_k+1})=\cdots=\nu(\alpha_{s_{k+1}-1})=m_k\]
with $m_0<m_1<\cdots<m_{r-1}.$ Here we ``redefine'' $s_\bullet$ and $r$ from earlier, but they will turn out to be what we need them to be. Regardless $s_0=0$ and $s_r=n,$ as required.

The key step is to think about the coefficients $a_\bullet$ as symmetric sums of the roots and then use the strong triangle inequality of our valuation. Explicitly, because $a_n=1,$ we get to say
\[\nu(a_k)=\nu\Bigg((-1)^{n-k}\sum_{\substack{S\subseteq[0,n)\\\#S=n-k}}\prod_{\ell\in S}\alpha_\ell\Bigg)\ge\min_{\substack{S\subseteq[0,n)\\\#S=n-k}}\nu\left(\prod_{\ell\in S}\alpha_\ell\right),\]
and because of the ultrametric triangle inequality, equality is achieved here if there is a subset $S$ which gives a product strictly smaller than all of the others. Anyways, this simplifies to
\[\nu(a_k)\ge\min_{\substack{S\subseteq[0,n)\\\#S=n-k}}\sum_{\ell\in S}\nu(\alpha_\ell)=\sum_{\ell=0}^{n-k-1}\nu(\alpha_\ell)\]
because we ordered the $\alpha_\bullet$ increasing with respect to $\nu.$ As mentioned earlier, equality here is achieved when $S$ is the definitive minimum, not just a minimum, which will only occur if we can't get substitute one of the $\nu(\alpha_\ell)$ with a different $\alpha_\bullet.$ Well, this only happens when we've used up an entire batch of $\alpha_\bullet$s with a particular $m_k,$ otherwise $\nu(\alpha_{n-k})=\nu(\alpha_{n-k+1}).$

Explicitly, equality is achieved when $k$ is one of $(n-s_0)$ (giving $0$), $(n-s_1)$ (with only $m_1$), $(n-s_2)$ (with only $m_1$ and $m_2$), and so on. On the Newton polygon, these points look like
\[\left(n-s_k,\sum_{\ell=0}^{s_k-1}\nu(\alpha_\ell)\right)=\left(n-s_k,\sum_{\ell=0}^{k-1}(s_{\ell+1}-s_\ell)m_\ell\right).\]
We claim that these define the lower convex hull of our Newton polygon. To see that this finishes the theorem, note that the displacement between consecutive points is
\[\left(n-s_k,\sum_{\ell=0}^{k-1}(s_{\ell+1}-s_\ell)m_\ell\right)-\left(n-s_{k+1},\sum_{\ell=0}^k(s_{\ell+1}-s_\ell)m_\ell\right)=(s_{k+1}-s_k,(s_{k+1}-s_k)(-m_k)),\]
which has horizontal length $s_{k+1}-s_k$ equal to the number of roots with $\nu(\alpha_\bullet)$ equal to the signed slope $-(-m_k)=m_k.$ This is what we wanted.

Quickly, it remains to show that these points are indeed the lower convex hull. Well, fix some $a_\bullet$ between $n-s_{k+1}$ and $n-s_k.$ We see
\[\nu(a_\bullet)\ge\sum_{\ell=0}^{s_k-1}\nu(\alpha_\ell)+\sum_{\ell=s_k}^{n-\bullet-1}\nu(\alpha_\ell)=\nu(a_{n-s_k})+(n-s_k-\bullet)m_\ell.\]
So the slope from $(n-s_k,\nu(a_{n-s_k}))$ to $(\bullet,\nu(a_\bullet))$ computes to
\[\frac{\nu(a_{n-s_k})-\nu(a_\bullet)}{n-s_k-\bullet}\le-m_\ell.\]
We recall that the equality is achieved for connecting $(n-s_k)$'s point to $(n-s_{k+1})$'s point; the inequality here indicates that $(\bullet,\nu(a_\bullet))$ is above this segment. (Observe that making $\nu(a_k)$ larger will make the discrepancy worse, implying $\nu(a_k)$ must be on the larger end.) Thus, these $(n-s_k,\nu(a_{n-s_k}))$ do form the lower convex hull.

\subsubsection{January 8th}
Today I learned the proof of the $abc$ conjecture for polynomials, and notably function fields. (I think---there might be a hole.) The $abc$ conjecture states that for given $\varepsilon>0,$ there are only finitely many triples of positive pairwise coprime positive integers $(a,b,c)$ satisfying $a+b=c$ and violating
\[c<\op{rad}(abc)^{1+\varepsilon}.\]
Here $\op{rad}$ is the product of the distinct prime factors. Anyways, we're saying that the above equality holds most of the time.

Fix $K$ a perfect field. The $abc$ conjecture for function fields is called Mason's theorem and says that for pairwise coprime polynomials $a,b,c\in K[t]$ not all of which have vanishing derivative satisfying $a+b=c,$ we always have
\[\max\{\deg(a),\deg(b),\deg(c)\}\le\deg(\op{rad}_K(abc))-1.\]
That is, our corresponding inequality holds strictly and always. Again, $\op{rad}$ refers to the product of the distinct irreducible factors, but I am writing $\op{rad}_K$ because $\deg(\op{rad}_K(abc))$ is poorly behaved. We are actually going to show that
\[\max\{\deg(a),\deg(b),\deg(c)\}\le\deg(\op{rad}_{\overline K}(abc))-1,\]
where here $\deg(\op{rad}_{\overline K}(abc))$ is counting distinct irreducible factors in $\overline K,$ which just counts the number of distinct roots of $abc.$ However,
\[\deg(\op{rad}_{\overline K}(p))=\deg(\op{rad}_K(p))\]
for $p\in K[t]$ by unique prime factorization. Namely, multiplicativity of $\op{rad}$ and additivity of $\deg$ means that it suffices to take $p$ an irreducible in $K$ after decomposing $p$ into irreducible parts over $K.$ Then the statement is saying that the degree of the irreducible is the number of distinct roots in the algebraic closure, which holds because our field is perfect, where no double roots are permitted for irreducibles.

Anyways, the theme is to pass into an algebraicly closed field, and once we remark that it suffices to show
\[\max\{\deg(a),\deg(b),\deg(c)\}\le\deg(\op{rad}_{\overline K}(abc))-1,\]
as above, we may just assert $a,b,c\in\overline K[t]$ and pretend that $K$ was algebraicly closed the entire time. (Aside from $\op{rad},$ the degrees aren't changing.) So without loss of generality, $K$ is algebraicly closed. Also, for symmetry, we replace the condition $a+b=c$ with $a+b+c=0,$ which changes no degrees and is therefore safe. So it suffices to show
\[\deg(c)\le\deg(\op{rad}_{\overline K}(abc))-1\]
by symmetry.

The way we access $\op{rad}$ algebraically is through the formal derivative. To exhibit this, consider a nonzero $p\in K[t].$ Factoring in the algebraicly closed $K$ looks like
\[p(t)=\prod_{k=1}^P(t-\alpha_k)^{\nu_k},\]
which gives $\op{rad}(p)(t)=\prod_k(t-\alpha_k).$ Now, we know that we can detect double-roots with the formal derivative, but we can do better than this, for actually what we know is that the formal derivative decreases multiplicities by $1.$ Namely, we see, by the product rule,
\[p'(t)=\sum_{\ell=1}^P\left(\frac d{dt}(t-\alpha_\ell)^{\nu_\ell}\right)\prod_{\substack{k=1\\k\ne\ell}}^P(t-\alpha_k)^{\nu_k}.\]
However, we can factor out many of these factors like
\[p'(t)=\left(\prod_{k=1}^P(t-\alpha_k)^{\nu_k-1}\right)\Bigg(\sum_{\ell=1}^Am_\ell\prod_{\substack{k=1\\k\ne\ell}}^P(t-\alpha_k)\Bigg).\]
In particular, we see that if we add in $\op{rad}(p)$ into the mix, we will have $p\mid p'\op{rad}(p),$ and even $p\mid(p,p')\op{rad}(p).$ The precise result we're going to need is the resulting statement
\[\deg(p)\le\deg((p,p'))+\deg(\op{rad}(p)).\tag{$*$}\label{eq:rad-der}\]
Note the degrees make sense with $p\ne0.$

Continuing, the main character in our story is the Wronskian. In our case, we will say $W(a,b)=ab'-ba'$ for $a,b\in K[t],$ but there is a more general theory here about detecting linear dependence of polynomials and their derivatives (which I do not currently understand). We start with some details. We claim that $a+b+c=0$ implies $W(a,b)=W(b,c)=W(c,a)=:W.$ By symmetry, it suffices to show $W(a,b)=W(b,c)$ (and then cycle the letters of the argument $(a,b,c)\mapsto(b,c,a)$), where this follows by writing
\[W(b,c)=bc'-cb'=b(-a-b)'-(-a-b)b'=(-a'b-bb')+(ab'+bb')=ab'-ba'=W(a,b).\]
As another detail, we claim $W\ne0,$ for $W=0$ will force $a'=b'=c'=0.$ Indeed, we can show $a'=0$ because $W=0$ implies $ab'=ba',$ so $a\mid a'$ because $\gcd(a,b)=1.$ From this, $a'=0$ would follow because $a'\ne0$ would force $\deg(a)>\deg(a'),$ which violates $a\mid a'.$

Now for the magic. Because $a,b,c$ are pairwise coprime, we remark $\op{rad}_{\overline K}(abc)=\op{rad}_{\overline K}(a)\op{rad}_{\overline K}(b)\op{rad}_{\overline K}(c).$ So we want to show
\[\deg(c)\le\deg(\op{rad}_{\overline K}(a))+\deg(\op{rad}_{\overline K}(b))+\deg(\op{rad}_{\overline K}(c))-1.\]
In order to make \hyperref[eq:rad-der]{$(*)$} appear, we add $\deg((a,a'))$ and friends to both sides so that it suffices to show (the stronger)
\[\deg((a,a'))+\deg((b,b'))+\deg((c,c'))+\deg(c)\le\deg(a)+\deg(b)+\deg(c)-1.\]
Cancelling the $\deg(c),$ we need
\[\deg((a,a'))+\deg((b,b'))+\deg((c,c'))\le\deg(a)+\deg(b)-1.\]
Now, we do have to deal with the asymmetry of not having $\deg(c)$ on the right-hand side, which is where the Wronskian enters. Because $\deg(p')\le\deg(p)-1,$ we see 
\[\deg(W)=\deg(ab'-ba')\le\max\{\deg(ab'),\deg(ba')\}=\deg(a)+\deg(b)-1.\]
Thus, it suffices to show
\[\deg((a,a'))+\deg((b,b'))+\deg((c,c'))\le\deg(W).\]
This is now symmetric.

We are almost done. Because $W\ne0,$ we may show what we need by showing
\[(a,a')(b,b')(c,c')\mid W.\]
However, of course $(a,a')\mid ab'-ba'=W.$ And with $a,b,c$ pairwise coprime, the above follows.

\subsubsection{January 9th}
Today I learned the proof of Fermat's last theorem for polynomials over a perfect field, as a corollary from yesterday's work; notably this holds for function fields. Fix $n$ a positive integer. Further, suppose $a,b,c\in K[t]$ for $K$ perfect and not all $a^n,b^n,c^n$ have vanishing derivative with $abc\ne0.$ Now given
\[a^n+b^n=c^n,\]
we want to show that $n<3.$ Without loss of generality, we take $a,b,c$ pairwise relatively prime, which follows from $\gcd(a,b,c)=1.$

Plugging into Mason's theorem, we remark that $\op{rad}_{\overline K}(a^nb^nc^n)=\op{rad}(abc),$ so we can say
\[n\max\{\deg(a),\deg(b),\deg(c)\}<\deg(\op{rad}(abc)).\]
Now, $\op{rad}(abc)\mid abc,$ and $abc$ is nonzero, so we may also say
\[\deg(\op{rad}(abc))\le\deg(abc)=\deg(a)+\deg(b)+\deg(c)\le3\max\{\deg(a),\deg(b),\deg(c)\}.\]
Combining this with the estimate from Mason's theorem requires $n<3,$ as desired.

As a side remark, what we proved does look like the normal Fermat's last theorem except for requiring not all of them to have vanishing derivative. On one hand, this is done to disallow constant polynomials, for which this would cover the real Fermat's last theorem.

But even ignoring this meta-argument, there are real dangers to watch out for. If our characteristic is $p>0,$ then for any polynomials $a,b\in K[t],$ we have that
\[a^p+b^p=(a+b)^p\]
by the Frobenius automorphism. This is a solution to Fermat's last theorem with $n=p,$ but we notice that all of $a^p,b^p,(a+b)^p$ have vanishing derivative, so this is ruled out by the hypotheses. To be clear, there are not such problems in characteristic $0,$ where vanishing derivative really does mean constant.

Also, there are infinitely many nontrivial solutions with $n=2,$ aside from characteristic $2$ where $n=2$ forces vanishing derivative. Recall the parameterization of Pythagorean triples by
\[\left(m^2-n^2\right)^2+(2mn)^2=\left(m^2+n^2\right)^2.\]
Because this is a purely algebraic fact, it'll hold for out polynomials in $K[t],$ so this infinite family will give lots of nontrivial solutions for $n=2,$ for any $m,n\in K[t].$ I think the usual proof that these give all primitive Pythagorean triples can also be ported over to here in $K[t],$ but I haven't worked through the details.

\subsubsection{January 10th}
Today I learned which elements of $\FF_q[[t]]^\times$ can be written as the sum of two squares for $q$ odd. The main idea is that Hensel-type lifting will imply that an element of $K[[t]]^\times$ is a square if and only if its constant term is square, for $K$ of characteristic not $2.$ (In characteristic $2,$ everything is a square, so this is uninteresting.) The forward direction is easy; if $a(t)\in K[[t]]^\times$ is a square, then
\[a(t)\equiv b(t)^2\pmod t\]
implies that $a(0)=b(0)^2,$ so our constant term is square.

The reverse direction is essentially Hensel lifting, as promised. Fix $a(t)=\sum_kx_kt^k\in K[[t]]^\times$ with $x_0=y_0^2$ in $K$; note $x_0\ne0$ implies $y_0\ne0.$ We construct $b(t)\in K[[t]]^\times$ inductively. We show that for any $n,$ we may construct $b_n(t):=\sum_{k=0}^ny_kt^k$ so that $b_{n+1}\equiv b_n\pmod{t^{n+1}}$ and
\[a(t)\equiv b_n^2\pmod{t^{n+1}}.\]
Sending $n\to\infty$ gives the result; I suppose a more rigorous proof would need some topology on $K[[t]]$ like that exists in $\ZZ_p,$ but we don't bother. Anyways, our base case is $b_0(t)=y_0,$ for which the conclusion holds by construction of $y_0.$

Now suppose we have $b_n$ so that we can construct $b_{n+1}\equiv b_n\pmod{t^{n+1}}$ satisfying the conclusion. Well, we write $b_{n+1}(t)=b_n(t)+y_{n+1}t^{n+1}$ for $y_{n+1}$ to be determined later, and the conclusion requires
\[a(t)\equiv b_{n+1}^2=\left(b_n(t)+y_{n+1}t^{n+1}\right)^2\pmod{t^{n+2}}.\]
Observe that $2n+2\ge n+2$ because $n\ge0,$ so this expands to
\[a(t)\equiv b_n^2+2y_{n+1}t^{n+1}\left(\sum_{k=0}^ny_kt^k\right)\equiv b_n^2+\left(2b_0y_0t^{n+1}\right)y_{n+1}\pmod{t^{n+2}}.\]
It follows
\[y_{n+1}\equiv\frac{a(t)-b_n^2}{2b_0y_0t^{n+1}}\pmod t\]
determines $y_{n+1}$ and completes the inductive step. This division is well-defined because all of $2,b_0,y_0\in K^\times$ are units, and $t^{n+1}$ divides the numerator by the inductive hypothesis.

Anyways, it remains to talk about sums of two squares in $\FF_q[[t]]^\times$ for $q$ odd. This is somewhat anticlimactic and roughly amounts to showing that every element of $\FF_q^\times$ can be written as the sum of two squares and then using the above machinery. Indeed, if $a\in \FF_q[[t]]^\times$ has $a(0)=b_0^2+c_0^2,$ then not both $b_0=c_0=0,$ and so $a-b_0^2\in\FF_q[[t]]^\times$ with square constant term, so it is itself a square. It follows
\[a=b_0^2+c^2,\]
which finishes, albeit in a boring way.

It remains to show that every element of $\FF_q$ can be written as the sum of two squares, which is pretty much combinatorics. Fix $x\in\FF_q,$ and observe that we want a square in
\[\left\{x-y^2:y\in\FF_q\right\}.\]
This set has $\frac{q+1}2$ total elements. Additionally, the squares in $\FF_q$ also have $\frac{q+1}2$ elements, for these sets to be disjoint, we would need at least $q+1$ elements between them. Of course, we only have $\#\FF_q=q$ elements to go around, so this is impossible.

Extending the result from $K[[t]]^\times$ to all of $K[[t]]$ doesn't look hard. Namely, I think we just have to specify that the term of lowest degree is even. Certainly if our lowest degree is even, we can just divide it out, apply the above, and then factor it into the square root. And conversely, being a square lets us track the term of lowest degree, which gets directly squared and will therefore have even degree after the squaring.

Extending this from $\FF_q[[t]]^\times$ to all of $\FF_q[[t]]$ looks a bit harrier. The problem is that cancellation might come in between our squares to create terms which are not trivially the sum of two squares, as we had above. I don't know how to resolve this issue.

\subsubsection{January 11th}
Today I learned about transporting functions. Fix $X:\UU$ and parameterize $A,B:X\to\UU$ so that we have a ``function'' type family $(A\to B)(x):\equiv A(x)\to B(x)$ by abusing notation. Then we can think about transporting $f:A(x_1)\to B(x_1)$ to some function in $A(x_2)\to B(x_2)$ along $p:x_1=x_2,$ which is classically denoted by
\[\op{transport}^{A\to B}(p)(f):A(x_2)\to B(x_2).\]
The point here is that there is a natural way to do this by taking $A(x_2)\to A(x_1)$ along $p^{-1},$ then taking $A(x_1)\to B(x_1)$ by $f,$ and then returning $B(x_1)\to B(x_2)$ along $p.$ Here's the commutative diagram.
\begin{center}
    \begin{tikzcd}
        A(x_2) \arrow[r, dashed]             & B(x_2)                 \\
        A(x_1) \arrow[r, "f"] \arrow[u, "p"] & B(x_1) \arrow[u, "p"']
    \end{tikzcd}
\end{center}
In symbols, we can follow the diagram to generate
\[\lambda a.\op{transport}^B(p)\left(f(\op{transport}^A(p^{-1})(a))\right):A(x_2)\to B(x_2).\]
The hope is that this is equal to $\op{transport}^{A\to B}(p)(f).$

Certainly these are equal pointwise. For fixed $a,$ we can show that
\[C(x_1,x_2,p)(f)(a)\equiv\op{transport}^{A\to B}(p)(f)(a)=\op{transport}^B(p)\left(f(\op{transport}^A(p^{-1})(a))\right).\]
is inhabited by path induction on $p.$ Indeed, path induction means that it suffices to inhabit $C(x,x,\refl_x),$ but in this case $\op{transport}^\bullet(\refl_x)\equiv\op{id}_\bullet$ by definition, so we need to inhabit
\[f(a)=f(a),\]
for which $\op{refl}_{f(a)}$ of course works.

However, it is not clear how to go from pointwise equality to full function equality, and the short of is that we can't. We can exhibit
\[(f=g)\to\prod_{a:A}f(a)=g(a)\]
by path induction on $p:f=g,$ for then we only have to exhibit $f(a)=f(a),$ where $\op{refl}_{f(a)}$ of course works. (Here we are working with arbitrary dependent functions $f,g:\prod_{a:A}B(a),$ but it works in this context.) It is an axiom, the ``function extensionality axiom,'' that this mapping is an equivalence, and this provides us a quasi-inverse
\[\op{funext}:\left(\prod_{a:A}f(a)=g(a)\right)\to(f=g).\]
Returning to the problem at hand, it follows that we can show the full equality
\[\op{transport}^{A\to B}(p)(f)=\lambda a.\op{transport}^B(p)\left(f(\op{transport}^A(p^{-1})(a))\right)\]
by using $\op{funext}$ on our witnesses for each $a:A(x_1).$ So things are good.

There is a corresponding equality for dependent functions in the more general case, but it is a bit annoying. I will write that it is
\[\lambda a_2.\op{transport}^{\widehat B}\left(\op{pair}^=\left(p^{-1},\op{refl}_{p_*^{-1}(a_2)}\right)\right)\left(f\left(\op{transport}^A(p)(a_2)\right)\right)\]
Here $\widehat B$ refers to the function type family taking $w:\sum_{(x:X)}A(x)$ to $B(\op{pr}_1w)(\op{pr}_2w).$ In particular, $B(x_\bullet)(a_\bullet)\equiv\widehat B((x_\bullet,a_\bullet)).$ The diagram is as follows; it's roughly the same. For brevity, we let $a_2$ be the input into the transported $f$ and $a_1:\equiv p^{-1}_*(a_2).$
\begin{center}
    \begin{tikzcd}
        A(x_2) \arrow[r, dashed]                     & \widehat B((x_2,a_2))                                       \\
        A(x_1) \arrow[r, "f"] \arrow[u, "p:x_1=x_2"] & \widehat B((x_1,a_1)) \arrow[u, "{p:(x_1,a_1)=(x_2,a_2)}"']
    \end{tikzcd}
\end{center}
All the garbage with $\op{pair}^=$ is really just trying to witness that $(x_1,a_1)=(x_2,a_2),$ and it's not terribly difficult (remembering how $\op{pair}^=$ works for $\sum$-types) to see that this type-checks.

\subsubsection{January 12th}
Today I learned about the univalence axiom. The core here is to establish what it means for two types $A$ and $B$ to be equal, with respect to a universe $\UU,$ but we don't really have an easy way to introduce such equalities. To be clear, we can exhibit a function
\[\op{idtoeqv}:(A=_\UU B)\to(A\simeq B).\]
Indeed, the main idea is to view $\op{id}_\UU:\UU\to\UU$ as a family of types indexed by $\UU$; what this gives us is the ability to think about
\[\op{transport}^{\op{id}_\UU}(p):A\to B\]
given our $p.$ We claim that this is an equivalence.

The easy way to see this as trying to exhibit a function
\[\prod_{(A,B:\UU)}(A=_\UU B)\to(A\simeq B),\]
for which path induction says that it suffices to take $A\equiv B$ and $p\equiv\refl_A$ so that our $\op{transport}^\bullet(p)\equiv\op{id}_\UU.$ But we know that $\op{id}_\UU$ is an equivalence with itself as a quasi-inverse, so we're done.

That is a bit cheap, however. More generally, for a type family $P:X\to\UU,$ we have that $p:x=y$ witnesses
\[\op{transport}^P(p):P(x)\to P(y),\]
which we claim is an equivalence $P(x)\to P(y).$ Taking $P\equiv\op{id}_\UU$ recovers the case that we're interested in. To show that this is an equivalence, it suffices to note that
\[\op{transport}^P(p^{-1}):P(y)\to P(x)\]
is its quasi-inverse. Indeed, because $\op{transport}$ commutes with path concatenation (proven with path induction), we can say that
\[\op{transport}^P(p^{-1})\circ\op{transport}^P(p)=\op{transport}^P(p\cdot p^{-1})\stackrel*=\op{transport}^P(\refl_x)=\op{id}_{P(x)}.\]
I guess some care must taken on $\stackrel*=$ because $\op{transport}^P$ is a dependent function, but $p\cdot p^{-1}$ and $\op{refl}_x$ are both of type $x=x,$ so we may restrict $\op{transport}^P$ to a function $(x=x)\to P(x)\to P(x).$ Then $\stackrel*=$ follows by applying $\op{transport}^P$ to both sides of $p^{-1}\cdot p=\refl_x.$ Anyways, similar holds in the opposite direction.

The point here is that
\[\op{idtoeqv}(p)\equiv\op{transport}^{\op{id}}(p)\]
suffices, where we are abusing $\op{transport}^{\op{id}}(p)$ to witness the entire equivalence in $A\simeq B.$ Now, the univalence axiom for a universe $\UU$ is the assertion that $\op{idtoeqv}$ is an equivalence for the universe $\UU$; in particular, we get a function
\[\op{ua}:(A\simeq B)\to(A=_\UU B)\]
which is the quasi-inverse of $\op{idtoeqv}.$ More or less, this is how we get to introduce equalities $A=B$---exhibit an equivalence $A\simeq B.$

Quickly, observe that this is the same pattern as with function extensionality. We could build a function
\[\op{happly}:(f=g)\to(f\sim g)\]
pretty easily by path induction. However, it's not clear how to construct the quasi-inverse of this function, so we have an axiom establishing it. It will turn out that univalence implies function extensionality, but I don't know why.

\subsubsection{January 13th}
Today I learned the outline of how to use class field theory to classify primes of the form $x^2+ny^2,$ where $n\equiv1\pmod4$ and is squarefree. To introduce in the algebraic number theory, we write this as
\[p=\left(x+y\sqrt{-n}\right)\left(x-y\sqrt{-n}\right),\]
so really we're saying that $p$ splits completely in $K:=\QQ(\sqrt{-n})$ into principal ideals. Notably, our rink of integers is $\mathcal O_K=\ZZ[\sqrt{-n}],$ so principal really does mean $\left(x+y\sqrt{-n}\right).$ We remark that this is an equivalence because if $p$ splits completely into principal ideals in $K,$ then we can write $p=(\alpha)(\beta)=(\alpha\beta),$ so $p^2=\op{Norm}(\alpha)\cdot\op{Norm}(\beta).$ Not both $\alpha$ and $\beta$ may be units, so we must have (say) $p=\op{Norm}(\alpha),$ which finishes. So we get
\[p=x^2+ny^2\iff p=\mf p\mf q,\,\text{ with }\mf p,\mf q\subseteq\mathcal O_K\text{ principal}.\]

It remains to classify primes $p$ which split completely into $\mf p\mf q$ in $\mathcal O_K$ where $\mf p$ and $\mf q$ are principal. Now we bring in the class field theory. Let $L$ be the Hilbert class field of $K,$ which has a number of remarkable properties. For example,
\[\mf p\text{ is principal}\iff\mf p\text{ splits completely in }L.\]
This is called the Principal Ideal Theorem. So we need the factors of $p$ to split completely in $L.$ And noting the multiplicativity of $e(\bullet/p)$ and $f(\bullet/p),$ we see that the factors of $p$ split completely if and only if $p$ itself splits completely, for we know that $p$ should split completely as a hypothesis already. It follows
\[p=x^2+ny^2\iff p\text{ splits completely in }L.\]

We can provide some aesthetically nicer conditions on splitting completely in $L.$ If we fix $L=K[\alpha]$ with $f(x)$ the minimal polynomial of our $\alpha,$ then for all but finitely many primes, we can be reassured the splitting of $p$ depends on the factorization of $f\pmod p.$ The end goal is to be able to have our condition be
\[p=x^2+ny^2\iff f(x)\equiv0\pmod p\text{ has a solution},\]
but we can't say that yet because of ramification. Well, another remarkable property of $L$ is that it is an unramified extension of $K,$ so it is enough to check that $p$ is unramified in $K.$ (It should be unramified because it should split completely, but we must check this in order to apply Dedekind-Kummer.) Ramification can be checked by checking $p\nmid\disc\mathcal O_K=-4n.$ So for all but finitely many primes,
\[p=x^2+ny^2\iff p\nmid4n\text{ and }f(x)\equiv0\pmod p\text{ has a solution}.\]
In particular, $f(x)\equiv0$ having a solution means that $p$ has a factor of inertial degree $1$ in $L,$ but then all factors have inertial degree $1,$ so $p$ does split completely.

\subsubsection{January 14th}
Today I learned a somewhat reliable way to compute $L(1,\chi)$ for characters $\chi$ (that actually doesn't use the multiplicative structure of $\chi$). Let's say that $\chi$ has period $N,$ and then we may say
\[L(1,\chi)=\sum_{n=1}^\infty\frac{\chi(n)}n.\]
The main idea is to use the generating functions idea to think about $\frac1n$ as $\int_0^1x^{n-1}\,dx.$ We would like to use this property to turn the above sum into an integral of a rational function, so we first group terms by period, giving
\[L(1,\chi)=\sum_{n=0}^\infty\left(\sum_{k=1}^N\frac{\chi(k)}{k+nN}\right)\]
using periodicity. Note that this grouping does not actually reorder the the summation of terms, but we're not going to worry much about convergence issues anyways. Now we introduce this as an integral, rewriting things as
\[L(1,\chi)=\sum_{n=0}^\infty\left(\sum_{k=1}^N\chi(k)\int_0^1x^{k+nN-1}\,dx\right)=\int_0^1\sum_{n=0}^\infty\left(\sum_{k=1}^N\chi(k)x^{k+nN-1}\right)\,dx.\]
However, we remark that we can actually split up the sum as
\[L(1,\chi)=\int_0^1\left(\sum_{n=0}^\infty x^{nN}\right)\left(\sum_{k=1}^N\chi(k)x^{k-1}\right)\,dx=\boxed{\int_0^1\frac1{1-x^N}\left(\sum_{k=1}^N\chi(k)x^{k-1}\right)\,dx},\]
which is what we wanted. This integral can sometimes be computed directly, but it can certainly be approximated; the core difficulty is the $1-x^N$ in the denominator, which is difficult in general to do partial fractions on.

Let's compute a sum with this. I'm not going to do examples of the analytic class number formula because the integral is actually quite hairy for nontrivial $N.$ So let's evaluate
\[1+\frac13-\frac15-\frac17+\frac19+\frac1{11}-\frac1{13}-\cdots.\]
This isn't a character, but it doesn't matter. Pushing this through the machinery, this becomes
\[\int_0^1\frac{1+x^2-x^4-x^6}{1-x^8}\,dx=\int_0^1\frac{\left(1-x^4\right)\left(1+x^2\right)}{\left(1-x^4\right)\left(1+x^4\right)}\,dx=\int_0^1\frac{1+x^2}{1+x^4}\,dx.\]
The standard way to integrate $\int\frac1{1+x^4}\,dx$ is by partial fractions because $1+x^4=\left(1+x^2\right)^2-\left(x\sqrt2\right)^2.$ In particular, we see
\[\frac{1+x^2}{1+x^4}=\frac{1/2}{1-x\sqrt2+x^2}+\frac{1/2}{1+x\sqrt2+x^2}.\]
Thus, it remains to integrate
\[\frac12\int_0^1\frac1{1-x\sqrt2+x^2}\,dx+\frac12\int_0^1\frac1{1+x\sqrt2+x^2}\,dx.\]
Completing the square in the denominator, we see $(x\sqrt2\pm1)^2=2x^2\pm2\sqrt2x+1,$ so the integral is
\[\int_0^1\frac1{1+(x\sqrt2-1)^2}\,dx+\int_0^1\frac1{1+(x\sqrt2+1)^2}\,dx.\]
This integral collapses to
\[\frac{\arctan(\sqrt2-1)+\arctan(\sqrt2+1)-\arctan(1)-\arctan(-1)}{\sqrt2}.\]
Of course, $\arctan$ is odd, so this is really
\[\frac{\arctan(\sqrt2-1)+\arctan(\sqrt2+1)}{\sqrt2}.\]
Using the normal tricks, we see $\arctan(\sqrt2\pm1)=\arg\left(1+(1\pm\sqrt2)i\right),$ so we want
\[\frac{\arg\left(1+(1+\sqrt2)i\right)\left(1+(1-\sqrt2)i\right)}{\sqrt2}=\frac{\arg(2i)}{\sqrt2}.\]
So our sum is $\boxed{\textstyle\frac{\pi}{2\sqrt2}}.$

\subsubsection{January 15th}
Today I learned another interpretation of the fact that the class number measures the failure of unique prime factorization of elements, from \href{https://math.stackexchange.com/questions/1673432/what-are-some-applications-of-chebotarev-density-theorem}{this post}. The typical statement is that $h_K=1$ is equivalent to unique prime factorization of elements, coming directly from unique prime factorization of ideals, but with $h_K>1,$ we hope that prime factorization breaks more. Anyways, I think the better mentality to have is that $h_K$ measures how much ideals fail to be principal, for which the class number formula asserting
\[\iota_C(t)\sim\iota_{C'}(t)\]
for any two ideal classes $C$ and $C'$ does the trick. Namely, ideals are evenly distributed over our ideal classes, so $\frac1{h_K}$ of our ideals are principal, and larger $h_K$ is equivalent to more ideals failing to be principal.

However, failure of unique prime factorization is really about the failure of prime ideals to be principal, which is somewhat more sophisticated. We have from the above that ideals themselves distribute evenly among the ideal classes, but this (directly) has little to do with the primes' distribution. However, with the help with some class field theory, this can be done.

For concreteness, fix $K/\QQ$ a number field with Hilbert class field $H.$ For primes $\mf p\subseteq\mathcal O_K,$ we have the equivalence
\[\mf p\text{ principal}\iff\mf p\text{ splits completely in }H\]
from the Principal ideal theorem. However, we know that $\frac1{[H:K]}=\frac1{\#\op{Gal}(H/K)}$ of the primes in $\mathcal O_K$ split completely in $H,$ a fact that can be seen in an elementary way (as we did a few months ago) or directly from Chebotarev: $\mf p$ splits completely if and only if
\[\left(\frac{H/K}{\mf p}\right)=\op{id}_{\op{Gal}(H/K)}\]
from theory around the Frobenius. This will occur with density $\frac1{\#\op{Gal}(H/K)}$ by Chebotarev. So it follows that $\frac1{\#\op{Gal}(H/K)}$ of primes in $\mathcal O_K$ are principal. To finish off, we remark that
\[\#\op{Gal}(H/K)=h_K\]
because the Galois group is in fact the class group. (As usual, I am citing a lot of class field theory without proof.) From this it follows that $\frac1{h_K}$ of all primes in $\mathcal O_K$ are principal, which is cute.

As an aside, it doesn't feel immediately clear that the distribution of prime ideals across ideal classes should be substantially more difficult than the distribution of primes across ideal classes. However, we can compare the difficulty of showing that integers are evenly distributed across modular classes---which is trivial---with showing the primes are evenly distributed across modular classes---which is notoriously hard. What follows isn't new, but it's exposition.

In both cases, the techniques used to show one statement are not easily applicable to the other, and amusingly, in both cases, enough machinery is still enough to make the statement on primes appear clear. I hope the above statement is clear, assuming enough class field theory---it's really ``just a computation.'' As for Dirichlet's theorem, we can remark that it's a trivial application of Chebotarev to $\QQ(\zeta_n),$ for
\[p\equiv a\pmod n\iff\left(\frac{\QQ(\zeta_n)/\QQ}p\right)=\left(\sigma:\zeta_n\mapsto\zeta_n^a\right)\in\op{Gal}(\QQ(\zeta_n)/\QQ),\]
the latter of which occurs with density $\frac1n$ from Chebotarev. This is, similarly, ``just a computation'' with enough machinery. Namely, we didn't even have to talk about $L$ functions, though they are hiding in the Chebotarev.

\subsubsection{January 16th}
Today I learned the Poisson summation formula, from \href{https://kconrad.math.uconn.edu/blurbs/grouptheory/charthy.pdf}{Keith Conrad}. Assume $f:\RR\to\CC$ is a sufficiently nice function, which is my way of hand-waving out convergence issues. For example, we assume that $f$ has a Fourier transform denoted
\[\hat f(y)=\int_\RR f(x)e^{-2\pi iy}\,dx.\]
Then the statement of the Poisson summation formula is that
\[\sum_{n\in\ZZ}f(n)=\sum_{n\in\ZZ}\hat f(n).\]
The existence of the left-hand sum is guaranteed somewhat from the existence of the Fourier transform, but I'm not sure how to guarantee the right-hand side. Frankly, it doesn't matter too much to me.

The main trick here is to make the left-hand summation periodic so that we can write it as a Fourier series. Explicitly, fix
\[F(x)=\sum_{n\in\ZZ}f(x+n)\]
so that $F(x)=F(x+1).$ It follows that we can write
\[F(x)=\sum_{n\in\ZZ}c_ne^{2\pi inx},\]
and our coefficients are
\[c_n=\left\langle F(x),e^{2\pi inx}\right\rangle=\int_0^1F(x)e^{-2\pi inx}\,dx.\]
Expanding out $F$ again, we see that
\[c_n=\int_0^1\left(\sum_{n\in\ZZ}f(x+n)\right)e^{-2\pi inx}\,dx=\sum_{n\in\ZZ}\int_0^1f(x+n)e^{2\pi inx}\,dx.\]
Sending $x+n\mapsto x$ collapses this down to
\[c_n=\sum_{n\in\ZZ}\int_n^{n+1}f(x)e^{2\pi inx}\,dx=\int_\RR f(x)e^{2\pi inx}\,dx,\]
which is just $c_n=\hat f(n).$ Bringing it all together, we have
\[\sum_{n\in\ZZ}f(x+n)=F(x)=\sum_{n\in\ZZ}c_ne^{2\pi inx}=\sum_{n\in\ZZ}\hat f(n)e^{2\pi inx}.\]
Taking $x=0$ in the above recovers the Poisson summation formula.

I don't yet know any applications of the Poisson summation formula, but I suppose I'm moving towards the functional equation for $\zeta$ sometime in my future. Roughly speaking, it would appear that there might be cases in which the Fourier transform is easier to understand or easier to bound (which isn't at all obvious to me), and this is why this is useful.

\subsubsection{January 17th}
Today I learned a strengthening of Bauer's theorem, from \href{https://kconrad.math.uconn.edu/blurbs/gradnumthy/chebappn.pdf}{Keith Conrad}. For $K/\QQ$ a number field and extensions $L_1/K$ and $L_2/K$ with Galois closure $M/K,$ we have $\op{Spl}(L_1/K)=\op{Spl}(L_2/K)$ (up to a set of primes of density $0$) if and only if $L_2\subseteq L_1$ and
\[\bigcup_\sigma\sigma\op{Gal}(L_1/K)\sigma^{-1}=\op{Gal}(M/L_2).\]
\todo{finish this}

\subsubsection{January 18th}
Today I learned an alternative explanation for the decimal expansion of fractions like $\frac1{9801},$ from \href{https://math.stackexchange.com/a/102684/869257}{here}. I'm used to evaluating the arithmetico-geometric series $\sum nx^{n+1}$ directly, but in fact it is in general true that
\[\left(\sum_{n=1}^\infty x^n\right)^2=\sum_{n=1}^\infty nx^{n+1}.\]
There are a variety of ways to see this, one of which is directly computing the right-hand side by setting $S=\sum_{n=1}^\infty nx^{n+1}$ and seeing
\[\frac1xS-S=\sum_{n=1}^\infty x^n=\frac x{1-x}.\]
This rearranges to $S=\left(\frac x{1-x}\right)^2,$ which is what we wanted after expanding the geometric series. However, presenting this as that single equation makes it much easier to say something like
\[\left(\sum_{n=1}^\infty x^n\right)^2=\sum_{a=1}^\infty\sum_{b=1}^\infty x^ax^b=\sum_{n=2}^\infty\left(\sum_{a+b=2}1\right)x^n.\]
This quickly turns into our desired $S$ after shifting the index of the sum.

Anyways, from this we see
\[\sum_{n=1}^\infty\frac n{100^{n+1}}=\left(\sum_{n=1}^\infty\frac1{100}\right)^2=\frac1{99^2}.\]
This is how $\frac1{9801}$ is expanded. As an aside, I guess I should explain why $98$ is skipped in the decimal expansion. Well, consider the following point in the sum:
\[\frac1{9801}=\frac0{100}+\frac1{100^2}+\cdots+\frac{97}{100^{98}}+\frac{98}{100^{99}}+\frac{99}{100^{100}}+\frac{100}{100^{101}}+\cdots.\]
The $100$ is in fact too big for its position, so its $100$ gets shifted over to the previous $99,$ collapsing things into
\[\frac1{9801}=\frac0{100}+\frac1{100^2}+\cdots+\frac{97}{100^{98}}+\frac{99}{100^{99}}+\frac{00}{100^{100}}+\frac{00}{100^{101}}+\cdots.\]
So in fact the $98$ is present, but it was turned into a $99$ by the carry, which shifted everything one over. In fact, all of the $100$s shift one over: the $100$ was turned into $00$ in the above, but the $101$ next to it makes the $00$ into $01.$ Then $101$ turns to $02,$ and so on. This pattern continues into the familiar decimal expansion
\[\frac1{9801}=0.\overline{000102\cdots969799}.\]
I think a formal proof would induct on the $100$s, but I don't feel like doing that now.

\subsubsection{January 19th}
Today I learn an application of the (truncated) Poisson summation formula to compute the quadratic Gauss sum. We begin by presenting the truncated Poisson summation formula. Fix $f:\RR\to\CC$ satisfying some boundedness conditions to give is a Fourier series. Dirichlet's theorem on representing functions by a Fourier series on $f:[0,1]\to\CC$ says that
\[\frac{f(0^-)+f(1^+)}2=\sum_{n=-\infty}^\infty\int_0^1f(x)e^{-2\pi inx}\,dx.\]
That is, our Fourier series represents the average of the limits approaching the point for a piecewise continuous function. I don't know the proof of this, but Fourier analysis is sad. Running the above with $f(x+k)$ for integers $k$ reveals
\[\frac{f(k^-)+f(k+1^+)}2=\sum_{n=-\infty}^\infty\int_0^1f(x+k)e^{-2\pi ix}\,dx=\sum_{n=-\infty}^\infty\int_k^{k+1}f(x)e^{-2\pi inx}\,dx,\]
so summing up to $m$ implies that
\[\frac12f(0)+f(1)+\cdots+f(m-1)+\frac12f(m)=\sum_{n=-\infty}^\infty\int_0^mf(x)e^{-2\pi inx}\,dx.\]
This is the truncated Poisson summation formula. We remark that adding the flipped sum with $m\mapsto-m$ and then sending $m\to\infty$ recovers the original Poisson summation formula.

This is a different proof than the one we had before, but it has the advantage of dealing with finite sum. This is important because the quadratic Gauss sum is also finite. In particular, taking $f(k)=e^{2\pi ik^2/m}$ makes the summation read
\[g_m=\sum_{k=0}^{m-1}e^{2\pi ik^2/m}=\frac12f(0)+f(1)+\cdots+f(m-1)+\frac12f(m).\]
It follows that
\[g_m=\sum_{n=-\infty}^\infty\int_0^me^{2\pi ix^2/m}e^{-2\pi inx}\,dx.\]
From here, the proof is computation. Completing the square, this is
\[g_m=\sum_{n=-\infty}^\infty e^{-2\pi imn^2/4}\int_0^me^{2\pi i(x-mn/2)^2/m}\,dx.\]
In particular, we may shift the integral to
\[g_m=\sum_{n=-\infty}^\infty e^{-2\pi imn^2/4}\int_{mn/2}^{m(n+2)/2}e^{2\pi ix^2/m}\,dx.\]
We would like to use the sum to collapse the integral into an infinite one, but this requires some care due to the scale factor $e^{-2\pi imn^2/4}.$ Well, for $n$ even, $mn^2/4$ is an integer, so the scale factor is just $1.$ Thus,
\[\sum_{\substack{n=-\infty\\n\text{ even}}}^\infty e^{-2\pi imn^2/4}\int_{mn/2}^{m(n+2)/2}e^{2\pi ix^2/m}\,dx=\int_{-\infty}^\infty e^{2\pi ix^2/m}\,dx.\]
And for $n$ odd, $n^2\equiv1\pmod4,$ so the scale factor comes out to $i^{-m}.$ Thus,
\[\sum_{\substack{n=-\infty\\n\text{ odd}}}^\infty e^{-2\pi imn^2/4}\int_{mn/2}^{m(n+2)/2}e^{2\pi ix^2/m}\,dx=i^{-m}\int_{-\infty}^\infty e^{2\pi ix^2/m}\,dx.\]
It follows that
\[g_m=\left(1+i^{-m}\right)\int_{-\infty}^\infty e^{2\pi ix^2/m}\,dx=\left(1+i^{-m}\right)\sqrt m\int_{-\infty}^\infty e^{2\pi ix^2}\,dx.\]
This integral is not easy to evaluate, but it isn't dependent on $m,$ so we can evaluate it by taking $m=1.$ Here $g_m=1,$ so the integral is $\frac1{1+i^{-1}}=\frac{1+i}2.$ Thus,
\[g_m=\frac{\left(1+i^{-m}\right)(1+i)}2\sqrt m.\]
Because $i^m$ only depends on $m\pmod4,$ we might as well do the casework, which gives
\[\boxed{g_m=\begin{cases}
    (1+i)\sqrt m & m\equiv0\pmod4, \\
    \sqrt m & m\equiv1\pmod4, \\
    0 & m\equiv2\pmod4, \\
    i\sqrt m & m\equiv3\pmod4.
\end{cases}}\]
This finishes the proof.

I would be remiss if I didn't prove quadratic reciprocity from this. As usual, for $p$ an odd prime and $a$ not divisible by $p,$ define
\[g_p(a)=\sum_{k=0}^{p-1}\left(\frac kp\right)e^{2\pi iak/p}.\]
Adding in $\sum e^{2\pi ian/p}=0$ to this sum, we find that
\[g_p(a)=\sum_{k=0}^{p-1}\left(1+\left(\frac kp\right)\right)e^{2\pi iak/p}=\sum_{k=0}^{p-1}e^{2\pi iak^2/p}\]
because $1+\left(\frac kp\right)$ is $2$ for nonzero squares, $1$ for zero, and $0$ otherwise. (Just count the number of times an element $k^2\pmod p$ appears in both sums.) In particular, $g_p(1)=g_p,$ which we evaluated above. However,
\[\left(\frac ap\right)g_p(a)=\sum_{k=0}^{p-1}\left(\frac{ak}p\right)e^{2\pi iak/p}=g_p(1).\]
To finish, we evaluate $g_{pq}$ twice, which is the main character of this story. One way is as above. Alternatively, we can use the Chinese remainder theorem to give a bijection $\ZZ/p\ZZ\times\ZZ/q\ZZ\to\ZZ/(pq)\ZZ$ by $(x,y)\mapsto xq+yp.$ (I haven't seen this before, but it is bijective.) It follows
\[g_{pq}=\sum_{x=0}^{p-1}\sum_{y=0}^{q-1}e^{2\pi i(xp+yq)^2/pq}.\]
We can expand this out because the cross term $2xy\cdot pq$ will die to periodicity. This is
\[g_{pq}=\left(\sum_{x=0}^{p-1}e^{2\pi ipx^2/q}\right)\left(\sum_{y=0}^{q-1}e^{2\pi iqy^2/p}\right)=g_q(p)g_p(q).\]
It follows that
\[\left(\frac pq\right)\left(\frac qp\right)=\frac{g_q(p)g_p(q)}{g_qg_p}=\frac{g_{pq}}{g_pg_q}.\]
Plugging in our evaluation of $g_\bullet$ from above recovers quadratic reciprocity.

\subsubsection{January 20th}
Today I learned an extension of the information-theoretic proof of the infinitude of primes, from \href{https://qchu.wordpress.com/2009/09/02/some-remarks-on-the-infinitude-of-primes/}{here}. The proof, roughly speaking, is that there aren't ``enough'' integers with all primes in a fixed set of $k$ primes named $p_1,\ldots,p_k.$ Essentially, if we can express all positive integers $n$ as
\[n=\prod_{\ell=1}^kp_\ell^{\nu_\ell},\]
then we can represent integers $n$ using the $k$-tuple $(\nu_1,\ldots,\nu_k),$ which only needs
\[\sum_{\ell=1}^k\log(\nu_\ell)<\log n\]
bits to represent, which is impossible.

To push this idea further, we fix a bound $n\le N$ and bound the number of integers $n$ with primes in $\{p_1,\ldots,p_k\}.$ In this case, fixing one such prime $p_\bullet,$ we see
\[2^{\nu_\bullet}\le p_\bullet^{\nu_\bullet}\le n=\prod_{\ell=1}^kp_\ell^{\nu_\ell}\le N.\]
It follows that $\nu_\bullet\le\log_2N$ for each $p_\bullet.$ Combining this for all $k$ primes, there are no more than $(\log_2N)^k$ total numbers which can be represented in this way. We remark that the inequality
\[(\log_2N)^k<N\]
for $N\to\infty$ is our proof of the infinitude of primes.

However, this provides a strange contrapositive of our statement. If we have any infinite sequence of positive integers $a_1,a_2,\ldots$ such that, for a bound $N,$ there are more than $(\log_2N)^k$ terms of the sequence less than $N,$ then our sequence represents more than $k$ primes. Quite simply, there are too many integers to be represented by any fixed set of $k$ primes. For concreteness, let's say our sequence is increasing so that we can set $N=a_n,$ from which we see
\[(\log_2a_n)^k<n\]
is enough to give more than $k$ primes. Rearranging, this bound is saying that
\[a_n<2^{\sqrt[k]n}\]
gives more than $k$ primes.

Adjusting for aesthetics, this implies that if an increasing sequence of positive integers $a_1,a_2,\ldots$ is $o\big(2^{\sqrt[k]n}\big)$ for any positive integer $k,$ then our sequence represents infinitely many primes. Because $2^{1/k}\to1^+$ as $k\to\infty,$ we might as well say that if our sequence of positive integers $a_1,a_2,\ldots$ is $o\left(\alpha^n\right)$ for any $\alpha>1,$ then our sequence represents infinitely many primes. Of course, the argument is a bit stronger than this, but this is aesthetically nice.

So, for example, $a_n=P(n)$ for any polynomial $P$ will suffice. This suggests that polynomials represent infinitely many primes, not because of any intrinsic property of polynomials, but because they grow too slowly. They give too many integers to represent the relatively sparse set of positive integers with only a fixed set of prime factors. This is somewhat remarkable because I'm used to doing the Euclidean proof by
\[a\equiv b\pmod p\implies P(a)\equiv P(b)\pmod p,\]
but this structure is technically unnecessary.

\subsubsection{January 21st}
Today I learned an alternative proof of Wilson's theorem, from group actions. The main result is something called the $p$-group fixed point theorem, from \href{https://qchu.wordpress.com/2013/07/09/the-p-group-fixed-point-theorem/}{here}. Namely, if a finite $p$-group $G$ acts on a set $X$ then the set of fixed points $X^G$ satisfies
\[|X|\equiv\left|X^G\right|\pmod p.\]
This follows directly from the class equation. Quickly, for $x\in G,$ we can biject cosets in $G/\op{Stab}_x$ with elements in the orbit $Gx$ (explicitly, $g\op{Stab}_x\mapsto gx$), we see that
\[|X|=\sum_{\text{orbits }Gx}|Gx|=\sum_{\text{orbits }Gx}[G:\op{Stab}_x]=\sum_{\text{orbits }Gx}\frac{|G|}{|\op{Stab}_x|}.\]
Now, $|\op{Stab}_x|$ divides $|G|$ and is therefore a power of $p,$ so $|G|/|\op{Stab}_x|$ is as well. So this is divisible by $p$ unless it is equal to $1,$ in which case $|Gx|=1$ exactly describes $x\in X^G,$ so $1\cdot\left|X^G\right|$ exactly measures the perturbations$\pmod p.$

Anyways, let's kill Wilson's theorem with our new black magic. We need to talk about $(p-1)!\pmod p,$ so we would like to make a set $X$ with size $(p-1)!.$ We can get a $p!$ by taking permutations of $\FF_p,$ but to get down to $(p-1)!,$ we need to divide out that factor of $p.$ The way we do this is by identifying the permutations
\[(a_1,\ldots,a_p)\sim(a_1+k,\ldots,a_p+k)\]
for any $(a_1,\ldots,a_p)\in S_p$ and $k\in\FF_p.$ Here the notation $(a_1,\ldots,a_p)$ denotes the permutation $n\mapsto a_n.$ So we define
\[X=S_p/\sim,\]
where $\sim$ is defined as above. One way to look at this is as fixing the last element: there is exactly one $k$ which will take $a_p=0,$ for example. However, we would like to make a $p$-group act on this thing, so having $p$-tuples makes it natural to make $G:=\ZZ/p\ZZ$ act on $X$ by shifts, like
\[[n]_p\cdot(a_1,\ldots,a_p)=(a_{1+n},\ldots,a_{p+n}),\]
where indices are taken$\pmod p.$

It remains to compute the number of fixed points of this action. Well, a permutation will be fixed as long as
\[(a_1,\ldots,a_p)\sim(a_{1+n},\ldots,a_{p+n})\]
for each $n\in\ZZ/p\ZZ.$ In fact, because $\ZZ/p\ZZ$ is cyclic, it really suffices to check $n=1,$ so we need a constant $k$ such that
\[(a_1,\ldots,a_p)=(a_2+k,\ldots,a_1+k,a_p+k).\]
That is, we need $a_\bullet\equiv a_{\bullet+1}+k\pmod p.$ To count the number of such tuples, we note that all tuples are equivalent to one with $a_1=0$ (explicitly, $k=-a_1$), in which case these look like $(0,k,2k,\ldots,(p-1)k).$ Because these tuples need to contain distinct elements, there are only $p-1$ options for $k.$ (All of these work because multiplication by $k$ is a bijection.) Thus, we have $\left|X^G\right|=p-1$ fixed points, which implies
\[(p-1)!\equiv(p-1)\equiv-1\pmod p.\]
This is Wilson's theorem, so we are done here.

\subsubsection{January 22nd}
Today I learned some representation theory, from Artin 10.1.1: all representations of finite groups with dimension $1$ have image that is finite cyclic. Unpacking the definitions, we are provided a homomorphism $\rho$ taking
\[\rho:G\to\op{GL}_1(\FF)\]
for some finite group $G$ and field $\FF.$ We need to show that $\rho(G)$ is finite cyclic.

The main step of the proof is to realize that $\op{GL}_1(\FF)\cong\FF^\times.$ For this we claim that
\[(T:\FF\to\FF)\stackrel\varphi\longmapsto T(1)\]
is an isomorphism. We run through the conditions.
\begin{itemize}
    \item Note $\varphi$ is well-defined because $T(1)\ne0,$ for this would imply $T(x)=xT(1)=0$ for any $x\in\FF,$ so $T$ would not be injective (or surjective).
    \item We know $\varphi$ is homomorphic by direct computation: $(T\circ T')(1)=T(T'(1))=T'(1)\cdot T(1).$
    \item Continuing, $\varphi$ is injective because $T(1)=T'(1)$ implies $T(x)=xT(1)=xT'(1)=T'(x)$ for any $x\in\FF,$ so $T=T'.$
    \item We see $\varphi$ is surjective because, for any $c\in\FF,$ $x\mapsto cx$ is in $\op{GL}_1(\FF)$ (this can be checked manually). This automorphism satisfies $1\mapsto c.$
\end{itemize}
It follows that $\varphi$ is an isomorphism, so we're done here.

From the above we see that we are really given a homomorphism $\rho:G\mapsto\FF^\times.$ Note that $\rho(G)$ is a finite subgroup of $\FF^\times,$ so we're really showing that finite subgroups of $\FF^\times$ are finite cyclic. With this transformed problem, we rename $\rho(G)$ to $G$ in order to be able to ignore $\rho.$ That is, $G\subseteq\FF^\times$ is a finite subgroup, and we want to show that $G$ is finite cyclic.

There are a couple of ways to finish from here. The cheapest is that Artin always assumes that our base field is $\FF=\CC,$ in which case any finite subgroup of $\CC^\times$ is finite cyclic by roots-of-unity type arguments. For example, take $n=|G|,$ and we can say that
\[G\subseteq\left\{x\in\FF^\times:x^n-1=0\right\}\]
by Lagrange's theorem on groups. In fact, this is an equality because the right-hand side has at most $n$ elements by Lagrange's theorem on polynomials. But in $\CC^\times,$ we know that the right-hand side is
\[\left\langle e^{2\pi i/n}\right\rangle.\]
This can be checked by, say, noting that both sets have size $n$ and satisfy $x^n-1=0.$

However, this can be done in significantly more generality, for what we want is true in any field. I might have seen this proof before, but I have since forgotten it. Anyways, we mimic the proof that finite fields have a multiplicative generator and borrow the finish from \href{https://math.stackexchange.com/a/335654/869257}{this post}. The main lemma we need is that for elements $g,h\in G,$ there exists an element whose multiplicative order which is the least common multiple of the orders of $g$ and $h.$

This is somewhat obnoxious to show. We begin by claiming that if $a:=\ord(g)$ and $b:=\ord(h)$ are coprime, then $\ord(gh)=ab.$ Indeed, certainly $(gh)^{ab}=1.$ And in the other direction, if $(gh)^x=1,$ then
\[1=(gh)^{ax}=h^{ax},\]
so $b\mid ax,$ and $b\mid x.$ Similarly, $a\mid x,$ so $ab\mid x.$ It follows $ab$ really is our order. Now, to finish our lemma, we remove the condition that $a$ and $b$ are coprime. However, we remark that we can separate the primes in $a$ and $b$ with unique prime factorization by writing
\[a_*=\prod_{\nu_p(a)\ge\nu_p(b)}p^{\nu_p(a)}\qquad\text{and}\qquad b_*=\prod_{\nu_p(b)>\nu_p(a)}p^{\nu_p(b)}.\]
Namely, $a_*b_*=\lcm(a,b)$ while having $\gcd(a_*,b_*)=1$ because no prime factors are shared. So we only need to remark that $g^{a/a_*}$ will have order $a_*$ and $h^{b/b_*}$ will have order $b_*$ in order to assert an element of order $a_*b_*,$ which is what we wanted.

To finish, we note that, because $G$ is finite, we may inductively say that there exists an element $g$ whose order $m$ is divisible by the order of all elements of $G.$ On one hand, $m\le|G|$ because $m$ is the order of some element of $G.$ On the other hand, every element of $G$ is the root of the polynomial
\[x^m-1=0,\]
so Lagrange's theorem on polynomials says that $|G|\le m.$ It follows that the order of $g$ is the full $m=|G|,$ so $G$ is finite cyclic. We remark here that this proof is really just the proof that every prime has a primitive root (via field theory), generalized.

\subsubsection{January 23rd}
Today I learned a proof of Maschkes Theorem: every representation of a finite group $\rho:G\to\op{GL}(V)$ can be decomposed into a direct sum of irreducible representations. (An irreducible representation is one that has no $G$-invariant subspaces.) The main claim is that $\rho$ preserves exists a positive-definite Hermitian form on $V$ preserved by $\rho(G)$; that is $\rho$ is always unitary for some form.

This is somewhat clever. We remark that certainly some positive-definite Hermitian form $\{\bullet,\bullet\}$ on $V$ exists---for example, choose an arbitrary basis to write $V\cong\CC^n$ for dimension $n,$ and then use the induced Hermitian form from $\CC^n.$ For this, we ``average'' $\{\bullet,\bullet\}$ over $G,$ by writing
\[\langle v,w\rangle=\frac1{|G|}\sum_{g\in G}\{\rho_gv,\rho_gw\}.\]
We claim that this $\langle\bullet,\bullet\rangle$ is a positive-definite Hermitian form that is also $G$-invariant.
\begin{itemize}
    \item This is positive-definite because
    \[\langle v,v\rangle=\frac1{|G|}\sum_{g\in G}\{\rho_gv,\rho_gv\}\ge0,\]
    inherited from the positive-definite nature of $\{\bullet,\bullet\}.$ Additionally, equality holds if and only if $v=0$ because $\rho_gv=0$ is equivalent to $v=0.$ (In particular, $\rho_g\in\op{GL}(V)$ is invertible.)
    \item This is bilinear because
    \[\langle au+bv,w\rangle=\frac1{|G|}\sum_{g\in G}\{a\rho_gv+b\rho_gv,\rho_gw\}=a\langle u,w\rangle+b\langle v,w\rangle\]
    after expanding out and distributing $\{\bullet,\bullet\}.$
    \item This is Hermitian because
    \[\langle w,v\rangle=\frac1{|G|}\sum_{g\in G}\{\rho_gw,\rho_gv\}=\overline{\langle v,w\rangle},\]
    again by inheriting $\{\bullet,\bullet\}$ being Hermitian.
    \item Finally, this is $G$-invariant because for any $h\in G,$ we have that $g\mapsto gh$ is a bijection over $G$ because of left-cancellation. Thus, we may write
    \[\langle\rho_hv,\rho_hw\rangle=\frac1{|G|}\sum_{g\in G}\{\rho_{gh}v,\rho_{gh}w\}=\frac1{|G|}\sum_{g\in G}\{\rho_gv,\rho_gw\}=\langle v,w\rangle\]
    by re-indexing the sum. We have used the fact that $\rho_\bullet$ is a homomorphism.
\end{itemize}
We remark that this means any finite subgroup of $\op{GL}(V)$---those that are the images of some representation---are made of unitary transformations, up to a change of basis.

Now that we know our representation gives unitary transformations, it's not terribly difficult to give our decomposition; equip $V$ with a fixed form $\langle\bullet,\bullet\rangle$ to make $\rho$ unitary. Decomposition is done inductively: the claim is that if $\rho$ has a $G$-invariant subspace named $W,$ then $\rho$ is the direct sum of its restriction to $W$ and its restriction to the orthogonal complement $W^\perp$ of $W.$ In this way, we can continually subdivide $\rho$ into direct sums until no parts have a $G$-invariant subspace.

Quickly, we recall that
\[W^\perp=\{v\in V:\langle v,w\rangle=0\text{ for all }w\in W\}.\]
It is true that $V=W\oplus W^\perp.$ For example, if we build an orthonormal basis $\{w_1,\ldots,w_m\}$ for $W$ (say, using Gram--Schmidt), then we can look at
\[v-\sum_{k=1}^m\langle v,w_k\rangle w_k\in W^\perp.\]
Indeed, for any basis vector $w_\ell,$ most of the dot products $\langle w_k,w_\ell\rangle$ are $0,$ unless $w_\ell=w_k,$ in which case $\langle v,w_\ell\rangle$ will properly cancel with it. A basis is probably unnecessary for this result, but whatever.

We note also that the restriction of $\rho$ to $W^\perp$ makes sense. Indeed, for $v\in W^\perp$ and any $w\in W,$ we have that
\[\langle\rho_\bullet v,w\rangle=\langle v,\rho_\bullet^{-1}w\rangle=0\]
because $\rho_\bullet^{-1}w\in W.$ Thus, $W^\perp$ is also $G$-invariant, so we may restrict $\rho$ to $W^\perp.$

Thus, because $V=W\oplus W^\perp,$ and $\rho$ restricts properly to each subspace, we may say that $\rho$ is the direct sum of its restrictions, say $\rho^\parallel$ to $W$ and $\rho^\perp$ to $W^\perp.$ To show how this works, we note that for any $v\in V,$ we can write $v=w^\parallel+w^\perp$ for $w^\parallel\in W$ and $w^\perp\in W^\perp,$ so
\[\rho_\bullet v=\rho_\bullet w^\parallel+\rho_\bullet w^\perp=\rho^\parallel_\bullet w^\parallel+\rho^\perp_\bullet w^\perp,\]
which is what we wanted. This completes the proof of Maschkes Theorem.

\subsubsection{January 24th}
Today I learned the Ping-pong lemma: if a group $G$ generated by $a$ and $b$ acts on a set $X$ has subsets $X_1$ and $X_2$ which do not contain each other but $a^\bullet X_1\subseteq X_2$ and $b^\bullet X_2\subseteq X_1$ (for nonzero $\bullet$), then $G=\langle a,b\rangle$ is free. This is not terribly interesting. Essentially, we need to show that any nontrivial word $g\in G$ is not the identity with the given action; visually, we are watching $X_1$ or $X_2$ ``bounce'' back and forth with each power of $a$ or $b.$

We can write any element $g\in G$ as $a^\bullet b^\bullet\cdots.$ Let's say that our element is
\[g_k=a^{n_1}b^{n_2}\cdots b^{n_{k-1}}a^{n_k}\]
for now; here $n_\bullet\ne0$ and $k$ is odd. In other words, $g$ alternates starting and ending with $a.$ We watch where this takes $X_1$ because that's all of the information we have now, so I guess we claim $g_kX_1$ by induction. If $k=1,$ this is the hypothesis. Otherwise, we note that
\[g_{k+2}X_1=g_kb^{n_{k+1}}a^{n_{k+2}}X_1\subseteq g_kb^{n_{k+1}}X_2\subseteq g_kX_1,\]
which is a subset of $X_2$ by the inductive hypothesis.

Now, having established the group action, we can quickly conclude that all of the $g_k$ aren't the identity. Indeed, this would imply
\[X_1=g_kX_1\subseteq X_2,\]
which is false by hypothesis.

The rest of the proof requires a trick that I had to look up. For any other element $g\in G,$ there aren't that many more forms for us to deal with, but they can all be reduced to the above by sufficient conjugation by $a.$ (Conjugation is the trick I missed.) For example, if $g$ looks like
\[g=b^{n_1}a^{n_2}\cdots a^{n_{k-1}}b^{n_k}\]
starting and ending with $b,$ then $aga^{-1}$ will start and end with $a$ and is therefore not the identity; it follows $g$ is also not the identity. Similarly, if $g$ looks like
\[g=a^{n_1}b^{n_2}\cdots a^{n_{k-1}}b^{n_k}\]
starting with $a$ and ending with $b,$ then $a^{n_1}ga^{-n_1}$ will start and end with $a$ and is therefore not the identity, so $g$ is again not the identity. Finally, if $g$ looks like
\[g=b^{n_1}a^{n_2}\cdots b^{n_{k-1}}a^{n_k}\]
starting with $b$ and ending with $a,$ then $g^{-1}$ starts with $a$ and ends with $b,$ so $g^{-1}$ and therefore $g$ are not the identity. This completes the proof that $G$ is freely generated by $a$ and $b.$

As an example, we show that the subgroup of $\op{SL}_2(\ZZ)$ generated by
\[A=\begin{bmatrix}1&2\\0&1\end{bmatrix},\qquad\text{and}\qquad B=\begin{bmatrix}1&0\\2&1\end{bmatrix}\]
is free. For this, we let $A$ and $B$ act on the upper-half plane $\mathcal H$ of $\CC$ in the usual way, and we let
\[X_1=\{z\in\mathcal H:|z|<1\},\qquad\text{and}\qquad X_2=\{z\in\mathcal H:|z|>1\}.\]
These sets are disjoint and therefore do not contain each other. It remains to check that $A^nX_1\subseteq X_2$ and $B^n\subseteq X_1$ for nonzero $n.$ To start, we can inductively show that
\[A^n=\begin{bmatrix}1&2n\\0&1\end{bmatrix},\qquad\text{and}\qquad B=\begin{bmatrix}1&0\\2n&1\end{bmatrix}.\]
So for $z\in X_1,$ we see
\[A^nz=\frac{z+2n}1=z+2n.\]
Namely, $A$ shifts $z$ right by $2n,$ and because the unit disk has diameter $2,$ every element in $X_1$ is displaced outside into $X_2.$ On the other hand, for $z\in X_2,$ we see
\[B^nz=\frac z{2nz+1}=\frac1{2n+1/z}.\]
Because $z\in X_2$ has magnitude larger than $1,$ $1/z$ has magnitude smaller than $1$ and will live in the unit disk. So by the same shifting argument with $A,$ $2n+1/z$ is back in $X_2,$ but then $B^nz$ forces us into $X_1$ again. This is the last condition we had to check, so we are done here.

\subsubsection{January 25th}
Today I learned about the regular representation of a group, from Artin as usual. The big-picture is that we can turn any group action into a representation: if $G$ acts on a set $S,$ then we we can index a basis of $\CC^{|S|}$ by $S$ named $e_s$ and then define $\rho_\bullet:G\to\op{GL}_{|S|}(\CC)$ by
\[\rho_ge_s=e_{gs}.\]
In other words, we just create the permutation matrix associated with the group action. As an aside, we remark that the character $\chi(g):=\op{trace}(\rho_g)$ associated with this action is fairly well-behaved. Using the given matrix representation for $\rho_g,$ we can directly compute the trace by summing along the diagonal, for each nonzero entry along the diagonal is $1$ and signifies the basis vector $e_s$ being fixed. Thus, we're counting elements fixed by $g,$ which is
\[\chi(g)=|\op{Stab}(g)|.\]
For example, $\chi(e)=|S|,$ which matches $\dim\CC^{|S|}$ as it should. (Namely, $\chi(e)=\dim V$ for any $\rho:G\to\op{GL}(V)$ because, after choosing any basis, $\rho_e$ turns into $I\in\op{GL}_\bullet(\CC).$)

Anyways, we talk about this construction in order to apply it to the action of $G$ on $G$ by left multiplication. Here, the character is especially simple. To compute $\chi(g),$ we remark that $h\mapsto gh$ has a fixed point $h=gh$ if and only if $g=e.$ Thus, we see
\[\chi(g)=\begin{cases}
    |G| & g=e, \\
    0 & g\ne e.
\end{cases}\]
Let's move towards proving that $|G|=\sum d_\bullet^2$ for $d_\bullet$ the dimensions of the irreducible characters of $G.$ Fix $\chi_1,\ldots,\chi_r$ as our nonisomorphic irreducible characters, where $d_\bullet=\dim\chi_\bullet.$

What we want to say is that the simple presentation of $\chi$ lets us write
\[\langle\chi,\chi_\bullet\rangle=\frac1{|G|}\sum_{g\in G}\chi(g)\overline{\chi_\bullet(g)}.\]
Because $\chi(g)\ne0$ only when $g=e,$ the sum collapses to $\chi(e)\chi_\bullet(e)=|G|d_\bullet,$ implying $\langle\chi,\chi_\bullet\rangle=d_\bullet.$ However, the $\chi_\bullet$ form an orthonormal basis of characters (coefficients in $\ZZ,$ I guess), which we show shortly from the orthogonality relations. Then, these dot product relations imply that
\[\chi=d_1\chi_1+\cdots+d_r\chi_r.\]
Plugging $e$ into this equation gives the desired result.

We show that the $\chi_\bullet$ give an orthonormal basis of characters, where our coefficients are taken $\ZZ.$ We assume the orthogonality relations, which is honestly where most of the work is. For any representation $\rho,$ we showed a few days ago (Maschkes Theorem) that we can write
\[\rho=n_1\rho_1\oplus\cdots\oplus n_r\rho_r\]
for distinct irreducible characters $\rho_\bullet$ and integers $n_\bullet.$ Taking traces, it follows that
\[\chi=n_1\chi_1+\cdots+n_r\chi_r\]
for $\chi_\bullet=\op{trace}(\rho_\bullet).$ Thus, all characters can be represented as a $\ZZ$-linear combination of the $\chi_\bullet,$ which is what we needed. We remark that this representation is in fact unique, for $\langle\chi,\chi_\bullet\rangle=n_1$ from the orthogonality relations. Thus, the coefficients in the above representation are forced equal to $\langle\chi,\chi_\bullet\rangle,$ giving our basis.

As an aside, we remark that the $\chi_\bullet$ form a basis of the class functions, which are the set of functions $G\to\CC$ which are defined only over the conjugacy classes of $G.$ Fix $r$ the number of conjugacy classes. Viewing class functions as $r$-tuples of complex numbers implies that the space of class functions has dimension $r.$ But this is the number of irreducible characters (not shown here), and because the irreducible characters are orthonormal (also not shown here), this means that they must form a basis.

\subsubsection{January 26th}
Today I learned a (contrived) example of a function which is both epic and monic but neither injective or surjective. The main idea is that epic and monic only work for injectivity and surjectivity when we have enough functions to witness elements, so to break this, we limit our functions.

Namely, we consider the following category with $2$ objects and $3$ morphisms.
\begin{center}
    \begin{tikzcd}
        {\{1,2\}} \arrow["{\operatorname{id}_{\{1,2\}}}"', loop, distance=2em, in=305, out=235] \arrow[rr, "f\,:\,x\mapsto3"] &  & {\{3,4\}} \arrow["{\operatorname{id}_{\{3,4\}}}"', loop, distance=2em, in=305, out=235]
    \end{tikzcd}
\end{center}
For convenience, we fix $X=\{1,2\}$ and $Y=\{3,4\}.$ We can verify manually that this is a category. We have $2$ objects and $3$ morphisms, two of which are the required identity mappings. Here is the function composition table; we use $-$ to signify that the domain does not match the codomain. (Left is left is input to $\circ$; top is right input.)
\[\begin{array}{c|ccc}
    \circ & \op{id}_X & \op{id}_Y & f \\\hline
    \op{id}_X & \op{id}_X & - & -\\
    \op{id}_Y & - & \op{id}_Y & f\\
    f & f & - & -
\end{array}\]
This associates because they're functions on sets. Alternatively, this is by brute force: the identity vacuously associates with every function by erasing the identity, and any string of compositions we can make in this category has at most one non-identity function because there is only one non-identity function, and it cannot compose with itself. And finally, the identity behaves with $f$ as seen in the table.

To show that $f$ is monic and epic, we again proceed by brute force. For monic, we are interested in maps $Z\to X$ for an object $Z$ in our category. However, $Z=Y$ doesn't matter because there are no maps in this direction, so we are forced into $Z=X,$ and our only morphism $X\to X$ is $\op{id}_X.$ So we merely have to check that
\[f\circ\op{id}_X=f\circ\op{id}_X\implies\op{id}_X=\op{id}_X,\]
which is true without having to look at $f.$

Epic is similar. We are interested in maps $Y\to Z$ for an object $Z.$ However, $Z=X$ doesn't matter because there are no maps in this direction, so we are forced into $Z=Y,$ and our only morphism $Y\to Y$ is $\op{id}_Y.$ So we merely have to check that
\[\op{id}_Y\circ f=\op{id}_Y\circ f\implies\op{id}_Y=\op{id}_Y,\]
which is again true without having to look at $f.$

It remains to show that $f$ is neither injective nor surjective. Well, $f$ is not surjective because no $x\in X$ gives $f(x)=d,$ and $f$ is not injective because $1\ne 2$ but $f(1)=3=f(2).$ This completes the proof.

I guess there are two follow-up questions to this example. The most obvious is if there are non-contrived examples. I suspect there are, but I haven't spent very long looking; I know that Vakil has some examples. Alternatively, we can ask if there is a simpler/more contrived example.

I don't think the category can get any simpler with $2$ objects, but it might be possible with just $1.$ The identity moprhism will be bijective, so we will need at least $2$ morphisms. The issue, now, is that the second morphism will be able to compose with itself, so some care is needed in checking, so maybe this is not as simple. For example, the following does not work.
\begin{center}
    \begin{tikzcd}
        {\{1,2\}} \arrow["f=\lambda x.1"', loop, distance=2em, in=125, out=55] \arrow["\operatorname{id}"', loop, distance=2em, in=305, out=235]
    \end{tikzcd}
\end{center}
Namely, $f^\bullet=f,$ breaking both monic and epic. In fact, no finite set will be able to work for our object because $\op{Mor}(S,S)$ is finite in \texttt{Set}, so $f^\bullet=f$ eventually.

\subsubsection{January 27th}
Today I learned another (more) contrived example of a function which is both monic and epic but neither injective nor surjective, but now with only $1$ object. The cost of using $1$ object is that we must have infinitely many morphisms: as discussed yesterday, we must have more than one morphism because the identity ``should'' be bijective, and adding a non-identity morphism $f$ adds $f^1,f^2,\ldots.$ If this (monoid structure) is finite, then $f^n=\op{id}$ for some $n,$ meaning that $f$ is not epic/monic.

So we should have infinitely many morphisms, and as such we're going to need an infinite set as our object. The following works.
\begin{center}
    \begin{tikzcd}
        \mathbb Z \arrow["f\,:\,n\mapsto n^2"', loop, distance=2em, in=125, out=55] \arrow["\operatorname{id}"', loop, distance=2em, in=305, out=235]
    \end{tikzcd}
\end{center}
Of course, we also have the induced morphisms $f^2,f^3,\ldots$ not shown in the above diagram, but these make up all of our morphisms. Quickly, this is a category: because all maps are $\ZZ\to\ZZ,$ our morphisms have well-defined composition, and this composition is associative because it is function composition. And the identity map fixes everyone, so $g\circ\op{id}=\op{id}\circ g=g$ for any $g\in\op{Mor}(\ZZ,\ZZ).$

We show now that $f$ is neither injective nor surjective. This is quick: $f(-1)=f(1)=1,$ so $f$ is not injective. And the image of $f$ is always nonnegative, so $f$ cannot be surjective.

It remains to show that $f$ is both epic and monic. By setting $f^0:=\op{id},$ we note that all of our morphisms have the form $f^\bullet,$ so the epic and monic conditions look like
\[\begin{cases}
    f^n\circ f=f^m\circ f\implies f^n=f^m, \\
    f\circ f^n=f\circ f^m\implies f^n=f^m.
\end{cases}\]
Both of these are saying $f^{n+1}=f^{m+1}$ implies $f^n=f^m.$ We remark that, for this, it's enough to say $f^n=f^m$ implies $n=m,$ for then
\[f^{n+1}=f^{m+1}\implies n+1=m+1\implies n=m\implies f^n=f^m.\]

To show that $f^n=f^m$ implies $n=m,$ we note that $f\left(2^k\right)=2^{2k},$ so $f^n(2)=2^{2^n}$ (by induction). So from $f^n=f^m,$ we have
\[2^{2^n}=f^n(2)=f^m(2)=2^{2^m}.\]
Taking logarithms, for example, we conclude $n=m.$ This completes the proof.

We close with some remarks. Because our category only has the maps $f^\bullet,$ the epic and monic conditions degenerate quickly to just showing that our category is the free monoid on one letter. Typically, having this structure means that we're invertible (and therefore bijective), but our construction somewhat bypasses invertibility to just say that the functions are all distinct. The exact function $f(n)=n^2$ is merely chosen to be kind of fast while not injective to make the end of the proof work properly; $f(n)=\floor{n/2}$ over $\NN\to\NN$ also works.

\subsubsection{January 28th}
Today I learned that the genus of the $n^\text{th}$ hypercube graph $Q_n$ is $(n-4)2^{n-3}+1$ for $n\ge2$ from \href{https://www.cambridge.org/core/journals/canadian-journal-of-mathematics/article/genus-of-the-ncube/C0E3405F0B3267A4792D0DB8CFC140D4}{here}. Quickly, the graph is $Q_n$ as
\[K_2\times K_2\times\cdots\times K_2,\]
where $\times$ is denoting the Cartesian product of graphs. We remark that assigning a bit to each of the $K_2$ lets us label the vertices of $Q_n$ with binary strings so that two vertices are adjacent if and only if they differ in exactly one bit. Here's $Q_3,$ the cube.
\begin{center}
    \begin{tikzcd}
    \texttt{000} \arrow[ddd, no head] \arrow[rrr, no head] \arrow[rd, no head] &                                                    &                                  & \texttt{001} \arrow[ld, no head] \arrow[ddd, no head] \\
                                                                               & \texttt{100} \arrow[r, no head] \arrow[d, no head] & \texttt{101} \arrow[d, no head]  &                                                       \\
                                                                               & \texttt{110} \arrow[r, no head]                    & \texttt{111} \arrow[rd, no head] &                                                       \\
    \texttt{010} \arrow[ru, no head] \arrow[rrr, no head]                      &                                                    &                                  & \texttt{011}                                         
    \end{tikzcd}
\end{center}

We prove the genus bound by lower-bounding with the Euler characteristic and upper-bounding by actually embedding $Q_n$ in a surface of the required genus. We begin with the lower bound. The definition of $Q_n$ quickly implies $V=2^n,$ and the handshake lemma implies $2E=nV$ because (with the binary representation) each vertex is adjacent to $n$ others, so $E=n2^{n-1}.$ 

Continuing, each face---in any representation---has degree at least $4,$ not $3$ because the graph is bipartite. Indeed, here's a coloring to show $Q_3$ is bipartite.
\begin{center}
    \begin{tikzcd}
    \texttt{000} \arrow[ddd, no head] \arrow[rrr, no head] \arrow[rd, no head] &                                                    &                                  & \color{red}\texttt{001} \arrow[ld, no head] \arrow[ddd, no head] \\
                                                                               & \color{red}\texttt{100} \arrow[r, no head] \arrow[d, no head] & \texttt{101} \arrow[d, no head]  &                                                       \\
                                                                               & \texttt{110} \arrow[r, no head]                    & \color{red}\texttt{111} \arrow[rd, no head] &                                                       \\
    \color{red}\texttt{010} \arrow[ru, no head] \arrow[rrr, no head]                      &                                                    &                                  & \texttt{011}                                         
    \end{tikzcd}
\end{center}
In general, a vertex is colored red if and only if its bit sum is odd: because adjacent vertices differ in exactly one bit, this will always flip the parity of the bit sum. Anyways, it follows that the sum of the degree of the faces is at least $4F,$ so the handshake lemma implies $F\le2E/4=n2^{n-2}.$

Plugging into the Euler characteristic formula, we see that
\[2-2g=V-E+F\le2^n-n2^{n-1}+n2^{n-2}.\]
The right-hand side is $(4-n)2^{n-2},$ and rearranging says $g\ge(n-4)2^{n-3}+1.$ In particular by this point we know that $Q_n$ is not planar for $n\ge4.$

We now prove the upper-bound to finish the proof. Let $g_n=(n-4)2^{n-3}+1$ so that we want to embed each $Q_n$ into a surface of $g_n.$ We do this inductively; note that the square $Q_2$ embeds into the plane of genus $g_2=0,$ which is our base case. So suppose we can embed $Q_n$ so that we want to embed $Q_{n+1}.$ Well, $Q_{n+1}$ is just two copies of $Q_n$ with edges between copied vertices, so we should probably spend $2g_n$ bridges embedding the $Q_n.$ This leaves
\[g_{n+1}-2g_n=(n-3)2^{n-2}+1-(n-4)2^{n-2}-2=2^{n-2}-1\]
bridges left over to place. We remark that we currently have the $Q_n$ on separate surfaces (effectively different components), which more or less gives us a bridge to spend for free: this bridge will just connect the components. So we have $2^{n-2}$ bridges to go around.

Before continuing, we motivate this discussion with the following embedding $Q_4$ into a torus.
\begin{center}
    \begin{asy}
        unitsize(1.3cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        // Q2
        p0000=(0,0); p0001=(1,0); p0010=(0,1); p0011=(1,1);
        p0100=(-0.5,-0.5); p0101=(1.5,-0.5); p0110=(-0.5,1.5); p0111=(1.5,1.5);
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q2'
        pair v = (3,0);
        p1000=v+(0,0); p1001=v+(1,0); p1010=v+(0,1); p1011=v+(1,1);
        p1100=v+(-0.5,-0.5); p1101=v+(1.5,-0.5); p1110=v+(-0.5,1.5); p1111=v+(1.5,1.5);
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // outer handles
        draw(p0111--p1110,dotted);
        draw(p0101--p1100,dotted);
        draw(p0110 -- p0110+(0.25,0.25) -- p1111+(-0.25,0.25) -- p1111, dotted);
        draw(p0100 -- p0100+(0.25,-0.25) -- p1101+(-0.25,-0.25) -- p1101, dotted);
        // bridge
        pair a=p0001+(-0.125,0.2), b=p1000+(0.125,0.2);
        pair c=p0011+(-0.125,-0.2), d=p1010+(0.125,-0.2);
        fill(a--b--d--c--cycle, lightgray);
        draw(a -- b, red);
        draw(c -- d, red);
        // (1-0.2-0.2)/5 = 0.12
        draw(p0001 -- p0001+(-0.32,0.2+0.1) -- p1000+(0.32,0.2+0.1) -- p1000, dotted);
        draw(p0011 -- p0011+(-0.32,-0.2-0.1) -- p1010+(0.32,-0.2-0.1) -- p1010, dotted);
        draw(p0000 -- p0000+(0.32,0.2+0.2) -- p1001+(-0.32,0.2+0.2) -- p1001, dotted);
        draw(p0010 -- p0010+(0.32,-0.2-0.2) -- p1011+(-0.32,-0.2-0.2) -- p1011, dotted);
    \end{asy}
\end{center}
Here, the right $Q_3$ is ``mirrored'' with respect to the left one, and the connecting edges between them are dotted. The hole in our surface is the bridge connecting the two inner squares.

We bring up this example to motivate the idea of placing bridges between faces in each copy of $Q_n$ in order to give the necessary edges a bridge. However, some care is required: we expect there to be a full $n2^{n-2}$ faces while we have only $2^{n-2}$ bridges, so we need to select the faces we puncture somewhat carefully. Namely, only $\frac1n$ of the faces are going to get a bridge. But this is fine! Each vertex has degree $n,$ so we expect to only need approximately $\frac1n$ of the faces to have a bridge to account for each vertex's needed edges.

It remains to actually name these faces, for which a technical trick is required. We would like all of our faces to be quadrilaterals (as in the Euler characteristic lower bound), so we need to select $2^{n-2}$ quadrilaterals from our $2^n$ vertices. (No vertex should be chosen twice because the above analysis is sharp.) Well, one way to do this is to just hope that the vertices labeled
\[\texttt{[bitstring]00},\quad\texttt{[bitstring]01},\quad\texttt{[bitstring]10},\quad\texttt{[bitstring]11}\]
all lie on one face. Indeed, the bit string will have $n-2$ bits, so this does give $2^{n-2}$ faces, and each vertex is included in one of these faces exactly once. Call these quadrilaterals ``nibble quadrilaterals.''

However, to make this proof work, we must alter our induction. We actually prove that it is possible to embed $Q_n$ into a surface of genus $g_n$ for which the nibble quadrilaterals are actually faces. Again, $Q_2$ is our base case. Our inductive step for $Q_{n+1}$ still begins by placing our two copies of $Q_n$ onto two surfaces, which leaves us with the $2^{n-2}$ bridges left to place from before.

We have done the necessary work to finish the proof now. Call one copy of $Q_n$ the $\texttt{0}$ copy and the other copy the $\texttt{1}$ to denote the prefix. Now connect each of the $2^{n-2}$ nibble face in the $\texttt{0}$ copy to the corresponding nibble face in the $\texttt{1}$ face by one of our $2^{n-2}$ bridges. Thus, each $\texttt{0}v\in Q_n$ lives on some nibble face, and the bridge allows an edge from $\texttt{0}v$ to $\texttt{1}v.$ (All edges can fit onto the bridge by flipping the $\texttt{1}$ copy as in the $Q_4$ embedding.) This completes the embedding of $Q_{n+1}.$

\subsubsection{January 29th}
Today I learned a proof (of the basic fact) that every natural number occurs as the genus of some graph. This is motivated by the synthesis of two approaches: on one hand, we can note that the genus of a graph is really the genus of the surface which the graph is a skeleton. On the other hand, we note the following embedding of $Q_4$ (from yesterday) onto a torus.
\begin{center}
    \begin{asy}
        unitsize(1.3cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        // Q2
        pair hslant = (0.2,0);
        real w = 1.5;
        real h = 0.5;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-1.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
    \end{asy}
\end{center}
This forms the skeleton of the torus, and it gives ideas on how to generalize this to more holes. Specifically, we consider the following sequence of graphs.
\begin{center}
    \begin{asy}
        unitsize(1cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        /* idiot graph */
        // Q2
        pair hslant = (0.2,0);
        real w = 1.5;
        real h = 0.5;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-1.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // thick handles
        draw(p0100--p1100);
        draw(p0110--p1110);
        // Q3
        draw(p0100--p0110);
        // Q3'
        draw(p1100--p1110);
        
        /* Q4 */
        pair shift = (1,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        
        /* double torus */
        shift = (4,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        shift = (3,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        
        /* far left dots */
        label("$\cdots$", (11.5,-0.5));
    \end{asy}
\end{center}
Call this sequence of graphs $T_0,T_1,T_2,\ldots.$ We claim that the genus of $T_n$ is $g_n=n.$

It should be somewhat clear from the diagrams given that the genus of $T_n$ is at least upper-bounded by $n,$ for we have provided an embedding. We will be more rigorous about this later (and do bridge-making, as with yesterday), but for now we will deal with upper-bounding the genus. This is done with the Euler characteristic. Consider the following ``module.''
\begin{center}
    \begin{asy}
        unitsize(1.3cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        // Q2
        pair hslant = (0.2,0);
        real w = 1.5;
        real h = 0.5;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-1.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // thick handles
        draw(p0000--p1000, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        fill(p1100--p1110--p0110--p0100--cycle, lightred);
        //fill(p1101--p1111--p0111--p0101--cycle, lightred);
        // dotted handles
        draw(p0000--p1000, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
    \end{asy}
\end{center}
Here the red face signifies that we have subtracted off the leftmost vertices and edges of the normal $Q_4.$ Then $T_n$ (formally) is made of a $T_0$ attached with $n$ of the above modules to the left. Note that the subtracted vertices and edges are (exactly) made up by the $T_0$ on the left end followed by the right side of the modules thereafter. The awkward definition of the module ensures no double-counting.

So to count vertices and edges, we merely note that the module has $V=\#V(Q_4)-4=12$ and $E=\#E(Q_4)-4=28.$ Thus, for $T_n,$ we have
\[V_n=\underbrace4_{T_0}+\underbrace{12n}_{\text{modules}},\qquad E_n=\underbrace4_{T_0}+\underbrace{28n}_{\text{modules}}\]
To bound faces, we remark that $Q_4$ is bipartite, so our module is bipartite. Additionally, we can attach modules together in a way that respects the coloring by flipping the colors. For example, here is $T_2$ colored.
\begin{center}
    \begin{asy}
        unitsize(1cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        pair shift;
        /* double torus */
        // Q2
        pair hslant = (0.2,0);
        real w = 1.5;
        real h = 0.5;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-1.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // coloring
        dot(p0000, blue); dot(p0001, red); dot(p0011, blue); dot(p0010, red);
        dot(p0100, red); dot(p0101, blue); dot(p0111, red); dot(p0110, blue);
        dot(p1000, red); dot(p1001, blue); dot(p1011, red); dot(p1010, blue);
        dot(p1100, blue); dot(p1101, red); dot(p1111, blue); dot(p1110, red);
        // again
        shift = (3,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // thick handles
        draw(p0000--p1000, gray); draw(p0100--p1100, gray);
        draw(p0001--p1001, gray); draw(p0101--p1101, gray);
        draw(p0010--p1010, gray); draw(p0110--p1110, gray);
        draw(p0011--p1011, gray); draw(p0111--p1111, gray);
        // cover
        fill(p0100--p0101--p0001--p0000--cycle, white);
        fill(p0100--p0110--p0010--p0000--cycle, white);
        fill(p0101--p0111--p0011--p0001--cycle, white);
        // dotted handles
        draw(p0000--p1000, dotted+gray); draw(p0100--p1100, dotted+gray);
        draw(p0001--p1001, dotted+gray); draw(p0101--p1101, dotted+gray);
        draw(p0010--p1010, dotted+gray); draw(p0110--p1110, dotted+gray);
        draw(p0011--p1011, dotted+gray); draw(p0111--p1111, dotted+gray);
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // coloring
        dot(p0000, red); dot(p0001, blue); dot(p0011, red); dot(p0010, blue);
        dot(p0100, blue); dot(p0101, red); dot(p0111, blue); dot(p0110, red);
        dot(p1000, blue); dot(p1001, red); dot(p1011, blue); dot(p1010, red);
        dot(p1100, red); dot(p1101, blue); dot(p1111, red); dot(p1110, blue);
    \end{asy}
\end{center}
This pattern continues; we do not bother being too rigorous here. Anyways, this means that faces have degree at least $4,$ so $4F_n\le2E_n,$ and $F_n\le2+14n$ by handshake. We remark that all faces are quadrilaterals in our embedding, so we expect this estimate to be safe.

It follows from the Euler characteristic that
\[2-2g_n=V_n-E_n+F_n\le(4+12n)-(4+28n)+(2+12n).\]
This rearranges to $g_n\ge n,$ which establishes the needed lower bound.

To close, we provide another way to see the embedding, which is essentially a generalization of the $Q_4$ embedding by building bridges. In words, we can build a bridge between each of the inner faces of the modules, and with $n$ modules, this makes $n$ bridges, completing the embedding. Visually, we begin by lying a $T_n$ flat, as such.
\begin{center}
    \begin{asy}
        unitsize(1cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        pair shift;
        /* double torus */
        // Q2
        pair hslant = (0,0);
        real w = 1;
        real h = 1;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-2.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // again
        shift = (2,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // labels
        label("$\cdots$", (4.3,0.5));
        label("$\cdots$", (4.3,0.5)+v);
    \end{asy}
\end{center}
Here we have repressed the edges from the top to the bottom because they are the problematic ones. In this representation, we will flip the bottom half to its bottom on top. This lets us connect the bottom of the top of $T_n$ with the top of the bottom, and vice versa, as follows.
\begin{center}
    \begin{asy}
        unitsize(1cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        pair shift;
        /* double torus */
        // Q2
        pair hslant = (0,0);
        real w = 1;
        real h = 1;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-2.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // handles
        draw(p0100--p1110, dotted); draw(p0101--p1111, dotted);
        draw(p0110-- p0110+(-.25,-.25) -- p1100+(-.25,.25) --p1100, dotted);
        draw(p0111-- p0111+(-.25,.25) -- p0110+(-.25,.25) -- p0110+(-.5,0) --
            p1100+(-.5,0) -- p1100+(-.25,-.25) -- p1101+(-.25,-.25) -- p1101, dotted);
        // again
        shift = (2,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // handles
        draw(p0101--p1111, dotted);
        draw(p0111-- p0111+(-.5,.5) -- p0110+(-.5,.5)-shift -- p0110+(-.75,.25)-shift --
            p1100+(-.75,-0.25)-shift -- p1100+(-.5,-.5)-shift -- p1101+(-.5,-.5) -- p1101,
            dotted);
        // labels
        label("$\cdots$", (4.3,0.5));
        label("$\cdots$", (4.3,0.5)+v);
    \end{asy}
\end{center}
Now we build a bridge between the corresponding inner faces of each module. This lets us connect the remaining vertices, as follows.
\begin{center}
    \begin{asy}
        unitsize(1cm);
        pair p0000, p0001, p0010, p0011;
        pair p0100, p0101, p0110, p0111;
        pair p1000, p1001, p1010, p1011;
        pair p1100, p1101, p1110, p1111;
        pair shift;
        /* double torus */
        // Q2
        pair hslant = (0,0);
        real w = 1;
        real h = 1;
        p0000=(0,0); p0001=(w,0); p0010=(0,h)+hslant; p0011=(w,h)+hslant;
        p0100=(-w/2,-h/2)-hslant/2; p0101=(3*w/2,-h/2)-hslant/2;
        p0110=(-w/2,3*h/2)+1.5*hslant; p0111=(3*w/2,3*h/2)+1.5*hslant;
        // Q2'
        pair v = (0,-2.5);
        p1000=v+p0000; p1001=v+p0001; p1010=v+p0010; p1011=v+p0011;
        p1100=v+p0100; p1101=v+p0101; p1110=v+p0110; p1111=v+p0111;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // handles
        draw(p0100--p1110, dotted); draw(p0101--p1111, dotted);
        draw(p0110-- p0110+(-.25,-.25) -- p1100+(-.25,.25) --p1100, dotted);
        draw(p0111-- p0111+(-.25,.25) -- p0110+(-.25,.25) -- p0110+(-.5,0) --
            p1100+(-.5,0) -- p1100+(-.25,-.25) -- p1101+(-.25,-.25) -- p1101, dotted);
        // bridge
        pair a=p0000+(0.2,0.125), b=p1010+(0.2,-0.125);
        pair c=p0001+(-0.2,0.125), d=p1011+(-0.2,-0.125);
        fill(a--b--d--c--cycle, lightgray);
        draw(a -- b, red);
        draw(c -- d, red);
        // (1-0.2-0.2)/5 = 0.12
        draw(p0000 -- p0000+(0.2+0.1,.32) -- p1010+(0.2+0.1,-.32) -- p1010, dotted);
        draw(p0001 -- p0001+(-0.2-0.1,.32) -- p1011+(-0.2-0.1,-.32) -- p1011, dotted);
        draw(p0010 -- p0010+(0.2+0.2,-.32) -- p1000+(0.2+0.2,.32) -- p1000, dotted);
        draw(p0011 -- p0011+(-0.2-0.2,-.32) -- p1001+(-0.2-0.2,.32) -- p1001, dotted);
        // again
        shift = (2,0);
        p0000 += shift; p0001 += shift; p0010 += shift; p0011 += shift;
        p0100 += shift; p0101 += shift; p0110 += shift; p0111 += shift;
        p1000 += shift; p1001 += shift; p1010 += shift; p1011 += shift;
        p1100 += shift; p1101 += shift; p1110 += shift; p1111 += shift;
        // Q3
        draw(p0000--p0001--p0011--p0010--cycle);
        draw(p0100--p0101--p0111--p0110--cycle);
        draw(p0000--p0100); draw(p0001--p0101);
        draw(p0010--p0110); draw(p0011--p0111);
        // Q3'
        draw(p1000--p1001--p1011--p1010--cycle);
        draw(p1100--p1101--p1111--p1110--cycle);
        draw(p1000--p1100); draw(p1001--p1101);
        draw(p1010--p1110); draw(p1011--p1111);
        // handles
        draw(p0101--p1111, dotted);
        draw(p0111-- p0111+(-.5,.5) -- p0110+(-.5,.5)-shift -- p0110+(-.75,.25)-shift --
            p1100+(-.75,-0.25)-shift -- p1100+(-.5,-.5)-shift -- p1101+(-.5,-.5) -- p1101,
            dotted);
        // bridge
        pair a=p0000+(0.2,0.125), b=p1010+(0.2,-0.125);
        pair c=p0001+(-0.2,0.125), d=p1011+(-0.2,-0.125);
        fill(a--b--d--c--cycle, lightgray);
        draw(a -- b, red);
        draw(c -- d, red);
        // (1-0.2-0.2)/5 = 0.12
        draw(p0000 -- p0000+(0.2+0.1,.32) -- p1010+(0.2+0.1,-.32) -- p1010, dotted);
        draw(p0001 -- p0001+(-0.2-0.1,.32) -- p1011+(-0.2-0.1,-.32) -- p1011, dotted);
        draw(p0010 -- p0010+(0.2+0.2,-.32) -- p1000+(0.2+0.2,.32) -- p1000, dotted);
        draw(p0011 -- p0011+(-0.2-0.2,-.32) -- p1001+(-0.2-0.2,.32) -- p1001, dotted);
        // labels
        label("$\cdots$", (4.3,0.5));
        label("$\cdots$", (4.3,0.5)+v);
    \end{asy}
\end{center}
This completes the embedding, which requires a total of $n$ bridges. Thus, $g_n\le n$ as well, so we conclude $g_n=n,$ which is what we wanted.

\subsubsection{January 30th}
Today I learned some stuff about the computability of convergence. To start, we note that given a computable sequence $(a_n)_{n=0}^\infty,$ it is not in general computable if
\[\sum_{n=0}^\infty a_n,\quad\text{or}\quad\prod_{n=0}^\infty a_n\]
converge. Indeed, we can reduce this quickly to halting. Given a program $\varphi,$ it is possible to compute
\[(a_\varphi)_n=\begin{cases}
    2 & \text{if }\varphi\text{ halts in }n\text{ steps}, \\
    0 & \text{otherwise}.
\end{cases}\]
Now, we do the casework on $\varphi$ halts.
\begin{itemize}
    \item If $\varphi$ halts in $N$ steps, then the infinite sum is $0$ after $N,$ so the sum is finite and converges. Similarly, this would mean that there is a $0$ in the infinite product, so the infinite product would converge to $0.$
    \item If $\varphi$ never halts, then the infinite sum sums $2$ for each $n,$ which of course diverges (say, by divergence test). And the infinite product is multiplying $2$ for each $n,$ which also diverges.
\end{itemize}
It follows that the infinite sum and product each converge if and only if $\varphi$ halts.

I find this somewhat unsatisfying because $(a_\varphi)_n$ is quite a contrived sequence. In reality, I'd like to think about this as a symbolic algebra problem: in the same way that it is uncomputable to determine if a given algebraic expression is identically $0,$ I feel like it should be uncomputable to determine if sequence converge. At a high level, I expect $\sum$ and $\prod$ to be able to look ``globally'' at a given function, allowing a weak $\forall$ check, which should be enough.

At the bare minimum, we should allow polynomials, which makes me feel like Hilbert's 10th will be involved. One basic difficulty is that the polynomials involved in Hilbert's 10th have lots of variables, but we are only summing over a single variable. This is not a huge problem, for good bijections $\NN\times\NN\to\NN$ exist, but it does require us to go outside of polynomials. For example,
\[f_2(n)=\left(\left|\floor{\sqrt n}-\left(n-\floor{\sqrt n}^2\right)\right|,n-\floor{\sqrt n}^2\right)\]
is a surjection $\NN\to\NN^2.$ (The idea here is that $\floor{\sqrt n}$ is a function slower than linear but still goes to infinity. Then we use it to dictate the coordinate sum for a contiguous set of inputs.) I don't like using $\sqrt\bullet$ because it is potentially undefined, but so it goes.

So because we have a function $f_2:\NN\to\NN^2,$ we can iteratively apply this to the first outputted coordinate to give a function $f_k:\NN\to\NN^k$ for any $k\in\NN.$ Thus, Hilbert's 10th begins to be possible here: at this point, we can do products. Indeed, fix a polynomial $P(x_1,\ldots,x_k)\in\ZZ[x_1,\ldots,x_k],$ and set
\[a_n=2|(P\circ f_k)(n)|.\]
Now, the infinite product $\prod_{n=0}^\infty a_n$ will get sent to $0$ if $P$ is $0$ anywhere in $\NN^k.$ If $P$ isn't $0$ anywhere, then each $a_\bullet\ge2,$ so the product will again not converge. It follows that determining convergence of infinite products even with just polynomials, $\floor\bullet,$ and $\sqrt\bullet$ is uncomputable, but for rather stupid reasons.

By this point, I think I've exhausted what easy arguments can give. For example, what made infinite products work is that products have a ``memory'' of previous elements of the sequence because $0$ ruins the party immediately. Infinite sums do not have this kind of luxury; we do remark that convergence of the infinite sum of
\[b_n=\prod_{k=0}^n|(P\circ f_k)(n)|\]
will be able detect if $P$ has a root because this forces a memory into the summation. However, I find this quite inelegant and unsatisfying.

\subsubsection{January 31st}
Today I learned an expression for Hermitian inner products in terms of the lengths they induce, from \href{http://estoyanov.net/files/MATAMATIKA/23982177-TheCauchy-Schwartz-Master-Class.pdf}{here}. These are, more or less, generalizations of the parallelogram law to work for Hermitian spaces.

To start, we fix a root of unity $\omega$ with $\omega^n=1$ but $\omega^2\ne1.$ Then we claim that, for $x,y\in\CC^\bullet,$
\[\langle x,y\rangle=\frac1n\sum_{k=0}^{n-1}\left\lVert x+\alpha^ky\right\rVert^2\alpha^k.\]
The feeling I get from this statement is that we are ``averaging'' the length over the various $\alpha^\bullet,$ in a way that looks analogous to the averaging process to define Hermitian inner products from group representations. I'm unsure if the relationship is more than aesthetic. Anyways, expanding out the length gives
\[\left\lVert x+\alpha^ky\right\rVert=\left\langle x+\alpha^ky,x+\alpha^k y\right\rangle=\langle x,x\rangle+\left\langle\alpha^ky,\alpha^ky\right\rangle+\left\langle\alpha^ky,x\right\rangle+\left\langle x,\alpha^ky\right\rangle.\]
Now, we dictate the fate of each of these terms as we sum over each $k.$
\begin{itemize}
    \item We note $\langle x,x\rangle$ is constant, so we have
    \[\langle x,x\rangle\sum_{k=0}^{n-1}\alpha^k=0\]
    because $\alpha\ne1$ is an $n$th root of unity.
    \item Because our inner product is Hermitian, $\left\langle\alpha^ky,\alpha^ky\right\rangle=\alpha^k\overline\alpha^k\langle y,y\rangle$ is constant as well (note $|\alpha|=1$), so this term vanishes as well.
    \item Again because our inner product is Hermitian, $\left\langle\alpha^ky,x\right\rangle=\alpha^k\langle y,x\rangle.$ Thus, we have
    \[\langle y,x\rangle\sum_{k=0}^{n-1}\alpha^{2k}=0\]
    because $\alpha^2\ne1$ is an $n$th root of unity. This is the only part of the proof which requires $\alpha^2\ne1.$
    \item Finally, $\left\langle x,\alpha^ky\right\rangle=\overline\alpha^k\langle x,y\rangle,$ so we have
    \[\langle x,y\rangle\sum_{k=0}^{n-1}\overline\alpha^k\alpha^k=n\langle x,y\rangle.\]
\end{itemize}
It follows that, after all of the dust has settled, we see
\[\frac1n\sum_{k=0}^{n-1}\left\lVert x+\alpha^ky\right\rVert^2\alpha^k=\frac nn\langle x,y\rangle=\langle x,y\rangle,\]
which is what we wanted.

What amuses me most about this result is that we can read it as a Riemann sum, in the same way that the discrete Fourier transform is a Riemann sum of the continuous Fourier transform. Indeed, take $n>4$ and $\alpha=e^{2\pi i/n},$ which tells us
\[\langle x,y\rangle=\frac1n\sum_{k=0}^{n-1}\big\lVert x+e^{2\pi i(k/n)}y\big\rVert^2e^{2\pi i(k/n)}.\]
Now, taking $n\to\infty$ and seeing this as a Riemann sum tells us
\[\langle x,y\rangle=\int_0^1\big\lVert x+e^{2\pi i\theta}y\big\rVert^2e^{2\pi i\theta}\,d\theta.\]
This is a bit prettier by taking $2\pi\theta\mapsto\theta,$ which gives
\[\langle x,y\rangle=\frac1{2\pi}\int_0^{2\pi}\left\lVert x+e^{i\theta}y\right\rVert^2e^{i\theta}\,d\theta.\]
This is cute, so we stop here.

We remark that we have not shown that the Riemann sum here is well-defined, but the induced norm is continuous, which is good enough. We do not show this continuity here, but I will outline: a Cauchy sequence $x_\bullet\to x$ by definition satisfies $\lVert x-x_\bullet\rVert\to0.$ But the triangle inequality gives
\[\left|\,\lVert x\rVert-\lVert x_\bullet\rVert\,\right|\le\lVert x-x_\bullet\rVert,\]
so $\lVert x_\bullet\rVert\to\lVert x\rVert$ as well, which is what we wanted.